---
title: "Final"
author: "Ricky Truong"
date: "May 2026"
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
  html_document:
    df_print: paged
geometry: margin=1in
fontsize: 11pt
---

# CO2: https://daac.ornl.gov/CMS/guides/CMS_DARTE_V2.html
# PM2.5: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/4GDRB1 

# Background

### Delete everything in environment

```{r, message = FALSE, warning = FALSE}
rm(list = ls())
```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(broom)
library(CausalArima)
library(colorspace)
library(ggnewscale)
library(ggpattern)
library(glue)
library(mgcv)
library(moderndive)
library(moments)
library(npreg)
library(plm)
library(plotly)
library(purrr)
library(scales)
library(scatterplot3d)
library(sf)
library(spdep)
library(splm)
library(tidyverse)
library(tigris)
library(tseries)
```

### Load data for county shapefile (from `library(tigris)`)

```{r}
# Load county shapefile (includes FIPS codes and geometry)
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  mutate(fips = paste0(STATEFP, COUNTYFP))

# Filter for contiguous U.S.
contiguous_fips <- setdiff(sprintf("%02d", 1:56),
                           c("02", "15", "60", "66", "69", "72", "78"))
counties_sf <- counties_sf %>%
  filter(STATEFP %in% contiguous_fips)
```

### Load data for crosswalk (between ZCTA and county)

```{r}
xwalk22 <- read_parquet("us_xwalks__census__zcta2county_yearly__2022.parquet")
xwalk21 <- read_parquet("us_xwalks__census__zcta2county_yearly__2021.parquet")
xwalk20 <- read_parquet("us_xwalks__census__zcta2county_yearly__2020.parquet")
xwalk19 <- read_parquet("us_xwalks__census__zcta2county_yearly__2019.parquet")
xwalk18 <- read_parquet("us_xwalks__census__zcta2county_yearly__2018.parquet")
xwalk17 <- read_parquet("us_xwalks__census__zcta2county_yearly__2017.parquet")
xwalk16 <- read_parquet("us_xwalks__census__zcta2county_yearly__2016.parquet")
xwalk15 <- read_parquet("us_xwalks__census__zcta2county_yearly__2015.parquet")
xwalk14 <- read_parquet("us_xwalks__census__zcta2county_yearly__2014.parquet")
xwalk13 <- read_parquet("us_xwalks__census__zcta2county_yearly__2013.parquet")
xwalk12 <- read_parquet("us_xwalks__census__zcta2county_yearly__2012.parquet")
xwalk11 <- read_parquet("us_xwalks__census__zcta2county_yearly__2011.parquet")
xwalk10 <- read_parquet("us_xwalks__census__zcta2county_yearly__2010.parquet")
xwalk09 <- read_parquet("us_xwalks__census__zcta2county_yearly__2009.parquet")
xwalk08 <- read_parquet("us_xwalks__census__zcta2county_yearly__2008.parquet")
xwalk07 <- read_parquet("us_xwalks__census__zcta2county_yearly__2007.parquet")
xwalk06 <- read_parquet("us_xwalks__census__zcta2county_yearly__2006.parquet")
xwalk05 <- read_parquet("us_xwalks__census__zcta2county_yearly__2005.parquet")
xwalk04 <- read_parquet("us_xwalks__census__zcta2county_yearly__2004.parquet")
xwalk03 <- read_parquet("us_xwalks__census__zcta2county_yearly__2003.parquet")
xwalk02 <- read_parquet("us_xwalks__census__zcta2county_yearly__2002.parquet")
xwalk01 <- read_parquet("us_xwalks__census__zcta2county_yearly__2001.parquet")
xwalk00 <- read_parquet("us_xwalks__census__zcta2county_yearly__2000.parquet")
```

### Load data for total CO2 emissions (at census block group level in kilograms)

```{r}
# Create path to the .gdb directory
co2_path <- "CMS_DARTE_V2_1735/data/DARTE_v2.gdb"

# List the available layers (feature classes) inside the .gdb
st_layers(dsn = co2_path)

# Load the specific layer
co2_cbg <- st_read(dsn = co2_path, layer = "DARTE_v2_blockgroup_kgco2_1980_2017")
```

### Load data for average PM2.5 concentration (at ZCTA level in micrograms per cubic meter)

```{r}
# Create path to the .parquet files
pm25_path <- "dataverse_files"

# List all .parquet files for years 2000–2020
file_list <- list.files(path = pm25_path,
                        pattern = "^pm25__ushap__zcta_yearly__20\\d{2}\\.parquet$",
                        full.names = TRUE)

# Read and combine all files into data frame
pm25_zcta <- file_list %>%
  lapply(read_parquet) %>%
  bind_rows()
```

# Data wrangling for `co2_cbg`

### Convert `co2_cbg` to tidy format

```{r}
co2_cbg <- co2_cbg %>%
  pivot_longer(cols = starts_with("kgco2_"),
               names_to = "year",
               names_prefix = "kgco2_",
               values_to = "value") %>%
  mutate(year = as.integer(year))
```

### Select for variables to avoid large data set

```{r}
co2_cbg <- co2_cbg %>%
  ungroup() %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  rename(geoid = GEOID) %>%
  select(geoid, year, value)
```

# Data wrangling for `xwalk`

### Combine everything into one crosswalk

```{r}
years <- 2000:2022
file_paths <- paste0("us_xwalks__census__zcta2county_yearly__", years, ".parquet")
xwalk_list <- map(file_paths, read_parquet)
xwalk <- bind_rows(xwalk_list)
```

# Aggregation for `pm25_zcta` (from ZCTA level to county level)

### Join `xwalk` and `pm25_zcta`

```{r}
pm25_county <- inner_join(xwalk, pm25_zcta, by = c("zcta", "year"))
```

### Aggregate from ZCTA level to county level

```{r}
pm25_county <- pm25_county %>% 
  group_by(year, county) %>% 
  mutate(weight = (arealand_pct / 100)) %>% 
  mutate(weight =  weight / sum(weight, na.rm = TRUE)) %>% 
  summarize(pm25 = sum(PM25 * weight, na.rm = TRUE),
            .groups = "drop")
```

# Aggregation for `co2_cbg` (from census block group level to county level)

### Create variable for county FIPS code

```{r}
co2_cbg <- co2_cbg %>%
  mutate(county = substring(geoid, 1, 5))
```

### Aggregate from census block group level to county level

```{r}
co2_county <- co2_cbg %>%
  group_by(year, county) %>% 
  summarize(co2 = sum(value, na.rm = TRUE),
            .groups = "drop")
```

# Finalizing files (`df` has CO2 and PM2.5 at county level from 2000 to 2017)

### Filter for contiguous U.S.

```{r}
pm25_county <- pm25_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)

co2_county <- co2_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

### Create `df` by joining `pm25_county` and `co2_county`

```{r}
df <- inner_join(pm25_county, co2_county, by = c("county", "year")) %>%
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

# Data visualization

### Aggregate CO2

```{r}
# Aggregate total CO2 for all counties by year
co2_total_year <- co2_county %>%
  group_by(year) %>%
  summarize(total_co2 = sum(co2, na.rm = TRUE))

# Check the result
head(co2_total_year)
```

### Plot aggregate CO2

```{r}
# Create graph of CO2 data
co2_plot <- co2_total_year %>%
  ggplot() +
  geom_line(aes(x = year, y = total_co2)) +
  geom_vline(aes(xintercept = 2005, linetype = "2005: Nonattainment Designation from NAAQS"), 
             color = "red") +
  labs(title = "Aggregated On-Road CO2 Emissions in the Contiguous U.S.",
       x = "Year",
       y = "CO2 (kilograms)",
       fill = "",
       linetype = "") +
  scale_linetype_manual(values = c("2005: Nonattainment Designation from NAAQS" = "dotted")) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Plot and save graph
co2_plot
ggsave("co2_plot.png",
       plot = co2_plot,
       bg = "white")
```

# Remove extreme counties (with CO2 2005-2010 relative changes above +/- 100%)

### Create `df_wide` (in wide format) from `df` (in long/tidy format)

```{r}
df_wide <- df %>%
  pivot_wider(names_from  = year,
              values_from = c(co2, pm25))
```

### Calculate 2005-2010 change in CO2 and PM2.5 for each county

```{r}
df_wide <- df_wide %>%
  mutate(co2_change = co2_2010 - co2_2005,
         co2_rel_change = co2_change / co2_2005,
         co2_rel_change_perc = co2_rel_change * 100,
         pm25_change = pm25_2010 - pm25_2005,
         pm25_rel_change = pm25_change / pm25_2005,
         pm25_rel_change_perc = pm25_rel_change * 100)
```

### Create vector for counties with extreme values of `co2_rel_change_perc`

```{r}
# Filter for extreme counties
extreme_counties <- df_wide %>% 
  filter(abs(co2_rel_change_perc) >= 100)

# Get vector of corresponding fips codes
extreme_fips <- extreme_counties$county
```

### Create `df_wide_filtered` and `co2_county_filtered` by removing extreme counties

```{r}
# Create df_wide_filtered
df_wide_filtered <- df_wide %>% 
  filter(!(county %in% extreme_fips))

# Create co2_county_filtered
co2_county_filtered <- co2_county %>% 
  filter(!(county %in% extreme_fips))
```

# Analysis with `CausalArima` library

<!-- ### Run `CausalArima()` on every available county to estimate statistics -->

<!-- ```{r, cache=TRUE} -->
<!-- # Set seed -->
<!-- set.seed(1) -->

<!-- # Initialize list to store results -->
<!-- results_list <- list() -->

<!-- # Get unique counties to avoid repeating counties over years -->
<!-- fips_vec <- unique(df_wide_filtered$county) -->

<!-- # Loop through each county -->
<!-- for (i in seq_along(fips_vec)) { -->

<!--   # Get county_df for current county -->
<!--   my_fips <- fips_vec[i] -->
<!--   county_df <- co2_county %>% -->
<!--     filter(county == my_fips) -->

<!--   # Filter for years less than or equal to 2010 (to avoid confounding) -->
<!--   county_df <- county_df %>% -->
<!--     filter(year <= 2010) -->

<!--   # Define intervention time point -->
<!--   intervention_date <- as.Date("2005-01-01") -->

<!--   # Create vector for dates -->
<!--   all_dates <- as.Date(paste0(county_df$year, "-01-01")) -->

<!--   # Create time series for outcome with yearly seasonality -->
<!--   y_ts <- ts(county_df$co2, frequency = 1) -->

<!--   # Use tryCatch to avoid full script crashing -->
<!--   tryCatch({ -->

<!--     # Run CausalArima() -->
<!--     ce <- CausalArima( -->
<!--       y = y_ts, -->
<!--       dates = all_dates, -->
<!--       int.date = intervention_date, -->
<!--       nboot = 1000 -->
<!--     ) -->

<!--     # Get ARIMA model specification -->
<!--     arma <- ce$model$arma -->
<!--     ar_order <- arma[1] -->
<!--     ma_order <- arma[2] -->
<!--     differencing_order <- arma[6] -->

<!--     # Get impact as list -->
<!--     imp <- impact(ce) -->

<!--     # Extract statistics via impact_boot -->
<!--     boot_list <- imp$impact_boot -->
<!--     boot_effect <- boot_list$effect_cum[3, ] -->
<!--     avg_effect <- imp$impact_boot$average$estimates[3] -->
<!--     cum_effect <- imp$impact_boot$effect_cum$estimates[3] -->
<!--     rel_effect <- imp$impact_boot$effect_cum$estimates[4] -->
<!--     p_value_one_sided <- as.numeric(boot_list$p_values["p"]) -->

<!--     # Store data as tibble -->
<!--     result_row <- tibble( -->
<!--       county = my_fips, -->
<!--       ar_order = ar_order, -->
<!--       differencing_order = differencing_order, -->
<!--       ma_order = ma_order, -->
<!--       avg_effect = avg_effect, -->
<!--       cum_effect = cum_effect, -->
<!--       rel_effect = rel_effect, -->
<!--       sd = boot_effect$sd, -->
<!--       ci_lower = boot_effect$inf, -->
<!--       ci_upper = boot_effect$sup, -->
<!--       p_value = min(1, 2 * p_value_one_sided) -->
<!--     ) -->

<!--     # Add to current results -->
<!--     results_list[[length(results_list) + 1]] <- result_row -->

<!--     # TEMPORARY CODE (DELETE LATER): Print progress -->
<!--     message(glue("{i}/{length(fips_vec)}: FIPS {my_fips} processed.")) -->

<!--   }, error = function(e) { -->
<!--     warning(glue("Error for FIPS {my_fips}: {e$message}")) -->
<!--   }) -->
<!-- } -->

<!-- # Save results -->
<!-- df_causal_arima <- bind_rows(results_list) -->
<!-- message("All counties processed. Final results saved.") -->
<!-- ``` -->

<!-- ### TEMPORARY CODE (DELETE LATER): Export `df_causal_arima` -->

<!-- ```{r} -->
<!-- write_csv(df_causal_arima, "~/df_causal_arima.csv") -->
<!-- ``` -->

### TEMPORARY CODE (DELETE LATER): Read `df_causal_arima`

```{r}
df_causal_arima <- read_csv("df_causal_arima.csv")
```

### Look at most common ARIMA model specifications

```{r}
df_causal_arima %>%
  count(ar_order, differencing_order, ma_order, sort = TRUE)
```

### Combine `df_wide_filtered` with `df_causal_arima` (both at county level)

```{r}
df_wide_filtered <- df_wide_filtered %>%
  left_join(df_causal_arima %>% select(county,
                                       avg_effect,
                                       cum_effect,
                                       rel_effect,
                                       sd,
                                       p_value),
            by = "county")
```

### Add polygons to `df_wide_filtered`

```{r}
# Join df_wide_filtered with shapefile
df_wide_filtered <- counties_sf %>% 
  inner_join(df_wide_filtered, by = c("fips" = "county")) 
```

### Plot `co2_rel_change_perc`

```{r}
# Define diverging color palette
div_colors <- c("#2166AC", "#4393C3", "#92C5DE", 
                "#F7F7F7",
                "#F4A582", "#D6604D", "#B2182B")

# Compute global limits for co2_rel_change_perc
global_min <- min(df_wide_filtered$co2_rel_change_perc, na.rm = TRUE)
global_max <- max(df_wide_filtered$co2_rel_change_perc, na.rm = TRUE)

# Create graph of differences
change_map <- ggplot(data = df_wide_filtered) + 
  geom_sf(aes(fill = co2_rel_change_perc), color = NA) +
  scale_fill_gradientn(colors = div_colors,
                       na.value = "white",
                       limits = c(global_min, global_max),
                       values = scales::rescale(c(global_min, 0, global_max)),
                       labels = label_number(accuracy = 1),
                       guide = guide_colorbar(barwidth = 20,
                                              barheight = 1,
                                              title.position = "top",
                                              title.vjust = 1)) + 
  labs(title = glue("Observed Change in CO2 Emissions by County, 2005-2010"),
       fill = "Change in CO2 (%)") +
  theme_minimal() +
  theme(legend.position = "bottom") +
  coord_sf(xlim = c(-125, -66), ylim = c(24, 50), expand = FALSE)

# Plot and save graph
change_map
ggsave("change_map.png",
       plot = change_map,
       width = 8, height = 6, dpi = 300,
       bg = "white")
```

### Plot `rel_effect_perc` and `pm25_rel_change_perc`

```{r}
# Create rel_effect_perc (i.e., rel_effect as a percentage)
df_wide_filtered <- df_wide_filtered %>% 
  mutate(rel_effect_perc = rel_effect * 100)

# Plot scatterplot with OLS line
scatterplot_ols <- ggplot(df_wide_filtered,
                          aes(y = rel_effect_perc,
                              x = pm25_rel_change_perc)) +
  geom_point(alpha = 0.25) +
  geom_smooth(method = "lm", se = TRUE, color = "#2166AC") +
  geom_hline(yintercept = 0, color = "grey", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "grey", linetype = "dashed") +
  labs(x = "PM2.5 Rel. Change from 2005 to 2010 (%)",
       y = "Estimated Rel. Effect of NAAQS on CO2 Emissions (%)",
       title = "PM2.5 Change and Effect on CO2 Emissions, 2005-2010: OLS",
       caption = "A negative x-value means PM2.5 decreased from 2005 to 2010 while
                  a negative y-value means observed CO2 was overall less than
                  counterfactual CO2 from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        plot.caption = element_text(size = 8),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        axis.text = element_text(size = 8))

# Plot and save graph
scatterplot_ols
ggsave("scatterplot_ols.png",
       plot = scatterplot_ols,
       width = 8, height = 6, dpi = 300,
       bg = "white")
```

### Model `rel_effect` and `pm25_rel_change` via OLS

```{r}
# Fit linear regression model
model_ols <- lm(rel_effect ~ pm25_rel_change, data = df_wide_filtered)
get_regression_table(model_ols)
```

### Model `rel_effect` and `pm25_rel_change` via GAM

```{r}
# Fit generalized additive model
model_gam <- gam(rel_effect ~ s(pm25_rel_change), data = df_wide_filtered)
summary(model_gam)

# Plot model
plot(model_gam)
```

# Analysis with `plm` library

### Prepare for `plm()` with `co2_county_filtered` (i.e., filtered)

```{r}
# Rename variable to fips in co2_county_filtered
co2_county_filtered <- co2_county_filtered %>%
  rename(fips = county)

# Filter for years less than or equal to 2004 (to fit to pre-treatment period)
co2_county_2004_filtered <- co2_county_filtered %>%
  filter(year <= 2004)

# Get shapefile of unique counties in co2_county_2004_filtered
fips_panel <- sort(unique(co2_county_2004_filtered$fips))
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  st_as_sf() %>%
  mutate(fips = str_pad(paste0(STATEFP, COUNTYFP), 5, "left", "0")) %>%
  filter(fips %in% fips_panel) %>%
  st_make_valid()

# Join co2_county_2004_filtered with shapefile (to add polygons/geometry)
co2_county_2004_filtered <- counties_sf %>%
  inner_join(co2_county_2004_filtered, by = c("fips" = "fips"))

# Create neighbor list using KNN to ensure each county has 4 neighbors
coords <- st_coordinates(st_centroid(counties_sf))
nb_knn <- knn2nb(knearneigh(coords, k = 4))

# Create listw object from nb_knn
listw_object <- nb2listw(nb_knn, style = "W")

# Ensure ordering matches plm()
regions <- counties_sf$fips
co2_county_2004_filtered <- co2_county_2004_filtered %>%
  filter(fips %in% regions) %>%
  mutate(fips = factor(fips, levels = regions)) %>%
  arrange(fips, year)

# Drop 1980 for every county (since there is no lag before)
df_one <- co2_county_2004_filtered %>%
  arrange(fips, year) %>%
  group_by(fips) %>%
  mutate(
    log_co2 = log(co2),
    log_lag_co2 = log(lag(co2))
  ) %>%
  filter(!is.na(log_lag_co2)) %>%
  ungroup()
```

### Run `plm()` with `df_one`: log-log model with 1 year of lag (i.e., most simple model)

```{r}
# Create panel dataframe from df_one
pdf <- pdata.frame(df_one, index = c("fips", "year"))

# Fit model
model_1 <- plm(
  log_co2 ~ log_lag_co2,
  data = pdf,
  model = "within",
  effect = "individual"
)
summary(model_1)

# Evaluate model with AIC
log_lik_val <- as.numeric(model_1$logLik)
k <- length(coef(model_1))
AIC_val_1 <- -2 * log_lik_val + 2 * k
AIC_val_1
```

This model is $\log(Y_{i, t}) = \beta_1 \log(y_{i, t - 1}) + \mu_{i} + \varepsilon_{i, t}$, which implies $\mathbb{E}[\log(Y_{i, t}) \mid \text{past}] = \beta_1 \log(y_{i, t - 1}) + \mu_{i}$. We estimate $\widehat{\mathbb{E}}[\log(Y_{i, t}) \mid \text{past}] = \hat{\beta}_1 \log(y_{i, t - 1}) + \hat{\mu}_{i}$. We thus want $\hat{\beta}_1$ (which we can extract from the model) and $\hat{\mu}_i$ (which we can calculate for each $i$).

Let $y_{i} = \log(Y_{i, t})$ and $x_{i} = \log(y_{i, t - 1})$. Our model is $y_{i} = \beta_1 x_{i} + \mu_{i} + \varepsilon_{i, t}$. Averaging over $t \in \{1, \ldots, T\}$, we have $\bar{y}_{i} = \beta_1 \bar{x}_{i} + \mu_{i} + \bar{\varepsilon}_{i, t}$. Since we assume $\mathbb{E}[\varepsilon_{i, t} \mid \text{past}] = 0$, for large $n$, $\bar{\varepsilon}_{i, t} \approx 0$, so $\bar{y}_{i} \approx \beta_1 \bar{x}_{i} + \mu_{i}$ $\implies$ $\mu_{i} \approx \bar{y}_{i} - \beta_1 \bar{x}_{i}$ $\implies$ $\hat{\mu}_{i} \approx \bar{y}_{i} - \hat{\beta}_{1} \bar{x}_{i}$.

### Use `model_1` to estimate counterfactual CO2 values

```{r}
# Drop geometry
df_one <- df_one %>%
  sf::st_drop_geometry()

# Get beta_1_hat
beta_hat <- coef(model_1)
beta_1_hat <- beta_hat["log_lag_co2"]

# Get mu_i_hat for each county
mu_hat_df <- df_one %>%
  group_by(fips) %>%
  summarize(
    mu_hat = mean(log_co2 - beta_1_hat * log_lag_co2, na.rm = TRUE),
    .groups = "drop"
  )

# Predict 2001 using observed 2000 CO2
pred_2001 <- df_one %>%
  filter(year == 2000) %>%
  left_join(mu_hat_df, by = "fips") %>%
  mutate(
    pred_year = 2001,
    pred_log_co2_2001 = beta_1_hat * log_co2 + mu_hat,
    pred_co2_2001 = exp(pred_log_co2_2001)
  ) %>%
  select(
    fips,
    pred_year,
    co2_2000 = co2,
    log_co2_2000 = log_co2,
    mu_hat,
    pred_log_co2_2001,
    pred_co2_2001
  )

# Display predicted vs. observed CO2 in 2001 for each county
df_one %>% 
  filter(year == 2001) %>% 
  rename(co2_2001 = co2) %>% 
  left_join(pred_2001, by = "fips") %>% 
  select(fips, co2_2000, pred_co2_2001, co2_2001)
```

### Run `plm()` with `df_three`: log-log model with 3 years of lag (i.e., more complicated model)

```{r}
# Drop 1980, 1981, 1982 for every county (since we're using three years of lag)
df_three <- co2_county_2004_filtered %>%
  arrange(fips, year) %>%
  group_by(fips) %>%
  mutate(
    log_co2 = log(co2),
    log_lag_co2_one = lag(log_co2, 1),
    log_lag_co2_two = lag(log_co2, 2),
    log_lag_co2_three = lag(log_co2, 3)
  ) %>%
  filter(
    !is.na(log_lag_co2_one),
    !is.na(log_lag_co2_two),
    !is.na(log_lag_co2_three)
  ) %>%
  ungroup()

# Create panel dataframe from df_three
pdf <- pdata.frame(df_three, index = c("fips", "year"))

# Fit model via plm()
model_2 <- plm(
  log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
  data = pdf,
  model = "within",
  effect = "individual"
)
summary(model_2)
```

This model is $\log(Y_{i, t}) = \beta_1 \log(y_{i, t - 1}) + \beta_2 \log(y_{i, t - 2}) + \beta_3 \log(y_{i, t - 3}) + \mu_{i} + \varepsilon_{i, t}$, which implies $\mathbb{E}[\log(Y_{i, t}) \mid \text{past}] = \beta_1 \log(y_{i, t - 1}) + \beta_2 \log(y_{i, t - 2}) + \beta_3 \log(y_{i, t - 3}) + \mu_{i}$. We estimate $\widehat{\mathbb{E}}[\log(Y_{i, t}) \mid \text{past}] = \hat{\beta}_1 \log(y_{i, t - 1}) + \hat{\beta}_2 \log(y_{i, t - 2}) + \hat{\beta}_3 \log(y_{i, t - 3}) + \hat{\mu}_{i}$. We thus want $\hat{\beta}_1$, $\hat{\beta}_2$, and $\hat{\beta}_3$ (which we can extract from the model) and $\hat{\mu}_i$ (which we can calculate for each $i$).

Let $y_{i, t} = \log(Y_{i, t})$ and $x_{i, t - j} = \log(y_{i, t - j})$. Our model is $y_{i, t} = \beta_1 x_{i, t - 1} + \beta_2 x_{i, t - 2} + \beta_3 x_{i, t - 3} + \mu_{i} + \varepsilon_{i, t}$. Averaging over $t \in \{1, \ldots, T\}$, we have $\bar{y}_{i} = \beta_1 \bar{x}_{i - 1} + \beta_2 \bar{x}_{i - 2} + \beta_3 \bar{x}_{i - 3} + \mu_{i} + \bar{\varepsilon}_{i, t}$. Since we assume $\mathbb{E}[\varepsilon_{i, t} \mid \text{past}] = 0$, for large $n$, $\bar{\varepsilon}_{i, t} \approx 0$, so $\bar{y}_{i} \approx \beta_1 \bar{x}_{i - 1} + \beta_2 \bar{x}_{i - 2} + \beta_3 \bar{x}_{i - 3} + \mu_{i}$ $\implies$ $\mu_{i} \approx \bar{y}_{i} - \beta_1 \bar{x}_{i - 1} - \beta_2 \bar{x}_{i - 2} - \beta_3 \bar{x}_{i - 3}$ $\implies$ $\hat{\mu}_{i} \approx \bar{y}_{i} - \hat{\beta}_{1} \bar{x}_{i - 1} - \hat{\beta}_{2} \bar{x}_{i - 2} - \hat{\beta}_{3} \bar{x}_{i - 3}$.

### Use `model_2` to estimate counterfactual CO2 values

```{r}
# Drop geometry
df_clean <- df_three %>%
  sf::st_drop_geometry()

# Get beta estimates
beta_hat <- coef(model_2)
beta_1_hat <- beta_hat["log_lag_co2_one"]
beta_2_hat <- beta_hat["log_lag_co2_two"]
beta_3_hat <- beta_hat["log_lag_co2_three"]

# Get mu_i_hat for each county
mu_hat_df <- df_clean %>%
  group_by(fips) %>%
  summarize(
    mu_hat = mean(
      log_co2 -
        beta_1_hat * log_lag_co2_one -
        beta_2_hat * log_lag_co2_two -
        beta_3_hat * log_lag_co2_three,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# Prepare full county-year data without geometry
df_all <- co2_county_filtered %>%
  sf::st_drop_geometry() %>%
  mutate(
    fips = as.character(fips),
    log_co2 = log(co2)
  ) %>%
  arrange(fips, year)

# Create wide data with observed log CO2 values
pred_wide <- df_all %>%
  filter(year >= 2002, year <= 2010) %>%
  select(fips, year, log_co2) %>%
  tidyr::pivot_wider(
    names_from = year,
    values_from = log_co2,
    names_prefix = "log_co2_"
  ) %>%
  left_join(mu_hat_df, by = "fips")

# For each year in 2005-2010, recursively predict log CO2
for (yr in 2005:2010) {
  
  # Define lag column names
  lag_1 <- paste0("log_co2_", yr - 1)
  lag_2 <- paste0("log_co2_", yr - 2)
  lag_3 <- paste0("log_co2_", yr - 3)
  pred_col <- paste0("pred_log_co2_", yr)
  
  # Predict log CO2 for current year
  pred_wide <- pred_wide %>%
    mutate(
      !!pred_col :=
        beta_1_hat * .data[[lag_1]] +
        beta_2_hat * .data[[lag_2]] +
        beta_3_hat * .data[[lag_3]] +
        mu_hat
    )
  
  # Replace log_co2_yr with predicted value so future years use recursive predictions
  pred_wide <- pred_wide %>%
    mutate(
      !!paste0("log_co2_", yr) := .data[[pred_col]]
    )
}

# Convert predictions to long format
pred_2005_2010 <- pred_wide %>%
  select(fips, starts_with("pred_log_co2_")) %>%
  tidyr::pivot_longer(
    cols = starts_with("pred_log_co2_"),
    names_to = "pred_year",
    values_to = "pred_log_co2"
  ) %>%
  mutate(
    pred_year = as.numeric(stringr::str_remove(pred_year, "pred_log_co2_")),
    pred_co2 = exp(pred_log_co2)
  )

# Join observed CO2 values for 2005-2010
pred_vs_obs <- df_all %>%
  filter(year >= 2005, year <= 2010) %>%
  select(fips, year, observed_co2 = co2, observed_log_co2 = log_co2) %>%
  left_join(
    pred_2005_2010,
    by = c("fips" = "fips", "year" = "pred_year")
  )

# Display predicted vs. observed CO2 in 2005-2010 for each county
pred_vs_obs %>%
  select(fips, year, observed_co2, pred_co2)
```

### Plot predicted vs. observed CO2 for individual county and for total (from `model_2`)

```{r}
# Plot for individual county
pred_vs_obs %>% 
  filter(fips == "01053") %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = observed_co2, color = "Observed CO2")) +
  geom_line(aes(y = pred_co2, color = "Predicted CO2")) +
  scale_color_manual(values = c("Observed CO2" = "blue", "Predicted CO2" = "red"))

# Plot for total/aggregated
pred_vs_obs %>% 
  group_by(year) %>% 
  summarize(total_observed_co2 = sum(observed_co2, na.rm = TRUE),
            total_pred_co2 = sum(pred_co2, na.rm = TRUE)) %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = total_observed_co2, color = "Total Observed CO2")) +
  geom_line(aes(y = total_pred_co2, color = "Total Predicted CO2")) +
  scale_color_manual(values = c("Total Observed CO2" = "blue", "Total Predicted CO2" = "red"))
```

### Analyze Normality of residuals (from `model_2`) for one county 

```{r}
# Define function to get residuals for a county
specified_fips <- "01053"
get_county_residuals_three_lag <- function(
    fips_use = specified_fips,
    co2_train = co2_county_2004_filtered
) {
  
  # Build 3-lag training data (pre-treatment only, up to 2004)
  df_three <- co2_train %>%
    sf::st_drop_geometry() %>%
    mutate(
      fips = as.character(fips),
      log_co2 = log(co2)
    ) %>%
    arrange(fips, year) %>%
    group_by(fips) %>%
    mutate(
      log_lag_co2_one = lag(log_co2, 1),
      log_lag_co2_two = lag(log_co2, 2),
      log_lag_co2_three = lag(log_co2, 3)
    ) %>%
    filter(
      !is.na(log_lag_co2_one),
      !is.na(log_lag_co2_two),
      !is.na(log_lag_co2_three)
    ) %>%
    ungroup()
  
  # Convert to panel dataframe
  pdf <- pdata.frame(df_three, index = c("fips", "year"))
  
  # Fit within / county fixed effects model
  model_2 <- plm(
    log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
    data = pdf,
    model = "within",
    effect = "individual"
  )
  
  # Extract beta estimates
  beta_hat <- coef(model_2)
  beta_1_hat <- beta_hat["log_lag_co2_one"]
  beta_2_hat <- beta_hat["log_lag_co2_two"]
  beta_3_hat <- beta_hat["log_lag_co2_three"]
  
  # Reconstruct county fixed effects mu_i
  mu_hat_df <- df_three %>%
    group_by(fips) %>%
    summarize(
      mu_hat = mean(
        log_co2 -
          beta_1_hat * log_lag_co2_one -
          beta_2_hat * log_lag_co2_two -
          beta_3_hat * log_lag_co2_three,
        na.rm = TRUE
      ),
      .groups = "drop"
    )
  
  # Get fitted values and residuals for county
  county_resid_df <- df_three %>%
    filter(fips == fips_use) %>%
    left_join(mu_hat_df, by = "fips") %>%
    mutate(
      fitted_log_co2 =
        beta_1_hat * log_lag_co2_one +
        beta_2_hat * log_lag_co2_two +
        beta_3_hat * log_lag_co2_three +
        mu_hat,
      
      residual_log_co2 = log_co2 - fitted_log_co2,
      
      fitted_co2 = exp(fitted_log_co2),
      residual_co2 = co2 - fitted_co2
    ) %>%
    select(
      fips, year, co2, log_co2,
      fitted_log_co2, residual_log_co2,
      fitted_co2, residual_co2,
      log_lag_co2_one, log_lag_co2_two, log_lag_co2_three,
      mu_hat
    )
  
  return(list(
    model = model_2,
    beta_hat = beta_hat,
    mu_hat_df = mu_hat_df,
    county_residuals = county_resid_df
  ))
}

# Get residuals for fips 01053
res_01053 <- get_county_residuals_three_lag(fips_use = "01053")

# Summarize residuals
res_01053$county_residuals
summary(res_01053$model)

# Plot histogram of residuals
resid_01053 <- res_01053$county_residuals$residual_log_co2
hist(resid_01053, breaks = 10, main = "Residuals for FIPS 01053",
     xlab = "Log-scale residuals")

# Plot Q-Q plot
qqnorm(resid_01053)
qqline(resid_01053)

# Run Shapiro-Wilk test (high p indicates Normality)
shapiro.test(resid_01053)

# Plot observed and fitted values
res_01053$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"), linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"), size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 01053",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

### Analyze Normality of residuals (from `model_2`) for all counties

```{r}
# Run on all counties
county_diag_df <- df_three %>%
  group_split(fips) %>%
  map_dfr(function(df_county) {
    
    # Get county residuals
    county_res <- df_county %>%
      left_join(mu_hat_df, by = "fips") %>%
      mutate(
        fitted =
          beta_1_hat * log_lag_co2_one +
          beta_2_hat * log_lag_co2_two +
          beta_3_hat * log_lag_co2_three +
          mu_hat,
        
        resid = log_co2 - fitted
      )
    
    r <- county_res$resid
    
    tibble(
      fips = unique(df_county$fips),
      n = length(r),
      mean_resid = mean(r),
      sd_resid = sd(r),
      skewness = moments::skewness(r),
      kurtosis = moments::kurtosis(r),
      shapiro_p = shapiro.test(r)$p.value,
      jb_p = jarque.bera.test(r)$p.value,
      acf1 = acf(r, plot = FALSE)$acf[2]
    )
  })

# Plot histogram of p-values from Shapiro-Wilk test
ggplot(county_diag_df, aes(x = shapiro_p)) +
  geom_histogram(bins = 30) +
  theme_minimal()

# Get proportion of counties that have low p-value (non-Normality)
mean(county_diag_df$shapiro_p < 0.01)

# Display counties with low p-value (non-Normality)
county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1:10)
```

### Analyze Normality of residuals (from `model_2`) for a "bad" county

```{r}
# Get fips and residuals of a county with low p-value (non-Normality)
bad_fips <- county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1) %>%
  pull(fips)
bad_res <- get_county_residuals_three_lag(
  fips_use = bad_fips
)$county_residuals

# Plot residuals over time
ggplot(bad_res, aes(x = year, y = residual_log_co2)) +
  geom_line() +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = paste("Residual Time Series:", bad_fips),
    x = "Year",
    y = "Residual"
  ) +
  theme_minimal()

# Plot histogram of residuals
ggplot(bad_res, aes(x = residual_log_co2)) +
  geom_histogram(bins = 10) +
  labs(
    title = paste("Residual Histogram:", bad_fips),
    x = "Residual"
  ) +
  theme_minimal()

# Plot Q-Q plot
ggplot(bad_res, aes(sample = residual_log_co2)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = paste("QQ Plot:", bad_fips)
  ) +
  theme_minimal()

# Plot observed and fitted values
res_29083 <- get_county_residuals_three_lag(fips_use = "29083")
resid_29083 <- res_29083$county_residuals$residual_log_co2
res_29083$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"),
            linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"),
             size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 29083",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

# Analysis with `splm` library

### Recreate `model_2` with `spml()` (i.e., three years of lag and spatial errors)

```{r}
# Create panel dataframe from df_three
pdf <- pdata.frame(df_three, index = c("fips", "year"))

# Fit model via spml()
model_3 <- spml(
  log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
  data = pdf,
  listw = listw_object,
  model = "within",
  effect = "individual"
)
summary(model_3)
```

```{r}
# Get estimated fixed effects
effects(model_3)
```

