---
title: "01: Prepare Data"
author: "Ricky Truong"
date: "Fall 2026"
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
  html_document:
    df_print: paged
geometry: margin=1in
fontsize: 11pt
---

# Background

### Delete everything in environment

```{r, message = FALSE, warning = FALSE}
rm(list = ls())
```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(broom)
library(CausalArima)
library(colorspace)
library(ggnewscale)
library(ggpattern)
library(glue)
library(mgcv)
library(moderndive)
library(moments)
library(npreg)
library(plm)
library(plotly)
library(purrr)
library(scales)
library(scatterplot3d)
library(sf)
library(spdep)
library(splm)
library(tidyverse)
library(tigris)
library(tseries)
```

### Load data for county shapefile (from `library(tigris)`)

```{r}
# Load county shapefile (includes FIPS codes and geometry)
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  mutate(fips = paste0(STATEFP, COUNTYFP))

# Filter for contiguous U.S.
contiguous_fips <- setdiff(sprintf("%02d", 1:56),
                           c("02", "15", "60", "66", "69", "72", "78"))
counties_sf <- counties_sf %>%
  filter(STATEFP %in% contiguous_fips)
```

### Load data for crosswalk (between ZCTA and county)

```{r}
xwalk22 <- read_parquet("us_xwalks__census__zcta2county_yearly__2022.parquet")
xwalk21 <- read_parquet("us_xwalks__census__zcta2county_yearly__2021.parquet")
xwalk20 <- read_parquet("us_xwalks__census__zcta2county_yearly__2020.parquet")
xwalk19 <- read_parquet("us_xwalks__census__zcta2county_yearly__2019.parquet")
xwalk18 <- read_parquet("us_xwalks__census__zcta2county_yearly__2018.parquet")
xwalk17 <- read_parquet("us_xwalks__census__zcta2county_yearly__2017.parquet")
xwalk16 <- read_parquet("us_xwalks__census__zcta2county_yearly__2016.parquet")
xwalk15 <- read_parquet("us_xwalks__census__zcta2county_yearly__2015.parquet")
xwalk14 <- read_parquet("us_xwalks__census__zcta2county_yearly__2014.parquet")
xwalk13 <- read_parquet("us_xwalks__census__zcta2county_yearly__2013.parquet")
xwalk12 <- read_parquet("us_xwalks__census__zcta2county_yearly__2012.parquet")
xwalk11 <- read_parquet("us_xwalks__census__zcta2county_yearly__2011.parquet")
xwalk10 <- read_parquet("us_xwalks__census__zcta2county_yearly__2010.parquet")
xwalk09 <- read_parquet("us_xwalks__census__zcta2county_yearly__2009.parquet")
xwalk08 <- read_parquet("us_xwalks__census__zcta2county_yearly__2008.parquet")
xwalk07 <- read_parquet("us_xwalks__census__zcta2county_yearly__2007.parquet")
xwalk06 <- read_parquet("us_xwalks__census__zcta2county_yearly__2006.parquet")
xwalk05 <- read_parquet("us_xwalks__census__zcta2county_yearly__2005.parquet")
xwalk04 <- read_parquet("us_xwalks__census__zcta2county_yearly__2004.parquet")
xwalk03 <- read_parquet("us_xwalks__census__zcta2county_yearly__2003.parquet")
xwalk02 <- read_parquet("us_xwalks__census__zcta2county_yearly__2002.parquet")
xwalk01 <- read_parquet("us_xwalks__census__zcta2county_yearly__2001.parquet")
xwalk00 <- read_parquet("us_xwalks__census__zcta2county_yearly__2000.parquet")
```

### Load data for total CO2 emissions (at census block group level in kilograms)

```{r}
# Create path to the .gdb directory
co2_path <- "CMS_DARTE_V2_1735/data/DARTE_v2.gdb"

# List the available layers (feature classes) inside the .gdb
st_layers(dsn = co2_path)

# Load the specific layer
co2_cbg <- st_read(dsn = co2_path, layer = "DARTE_v2_blockgroup_kgco2_1980_2017")
```

### Load data for average PM2.5 concentration (at ZCTA level in micrograms per cubic meter)

```{r}
# Create path to the .parquet files
pm25_path <- "dataverse_files"

# List all .parquet files for years 2000–2020
file_list <- list.files(path = pm25_path,
                        pattern = "^pm25__ushap__zcta_yearly__20\\d{2}\\.parquet$",
                        full.names = TRUE)

# Read and combine all files into data frame
pm25_zcta <- file_list %>%
  lapply(read_parquet) %>%
  bind_rows()
```

# Data wrangling for `co2_cbg`

### Convert `co2_cbg` to tidy format

```{r}
co2_cbg <- co2_cbg %>%
  pivot_longer(cols = starts_with("kgco2_"),
               names_to = "year",
               names_prefix = "kgco2_",
               values_to = "value") %>%
  mutate(year = as.integer(year))
```

### Select for variables to avoid large data set

```{r}
co2_cbg <- co2_cbg %>%
  ungroup() %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  rename(geoid = GEOID) %>%
  select(geoid, year, value)
```

# Data wrangling for `xwalk`

### Combine everything into one crosswalk

```{r}
years <- 2000:2022
file_paths <- paste0("us_xwalks__census__zcta2county_yearly__", years, ".parquet")
xwalk_list <- map(file_paths, read_parquet)
xwalk <- bind_rows(xwalk_list)
```

# Aggregation for `pm25_zcta` (from ZCTA level to county level)

### Join `xwalk` and `pm25_zcta`

```{r}
pm25_county <- inner_join(xwalk, pm25_zcta, by = c("zcta", "year"))
```

### Aggregate from ZCTA level to county level

```{r}
pm25_county <- pm25_county %>% 
  group_by(year, county) %>% 
  mutate(weight = (arealand_pct / 100)) %>% 
  mutate(weight =  weight / sum(weight, na.rm = TRUE)) %>% 
  summarize(pm25 = sum(PM25 * weight, na.rm = TRUE),
            .groups = "drop")
```

# Aggregation for `co2_cbg` (from census block group level to county level)

### Create variable for county FIPS code

```{r}
co2_cbg <- co2_cbg %>%
  mutate(county = substring(geoid, 1, 5))
```

### Aggregate from census block group level to county level

```{r}
co2_county <- co2_cbg %>%
  group_by(year, county) %>% 
  summarize(co2 = sum(value, na.rm = TRUE),
            .groups = "drop")
```

# Finalizing files (`df` has CO2 and PM2.5 at county level from 2000 to 2017)

### Filter for contiguous U.S.

```{r}
pm25_county <- pm25_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)

co2_county <- co2_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

### Create `df` by joining `pm25_county` and `co2_county`

```{r}
df <- inner_join(pm25_county, co2_county, by = c("county", "year")) %>%
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

# Remove extreme counties (with CO2 2005-2010 relative changes above +/- 100%)

### Create `df_wide` (in wide format) from `df` (in long/tidy format)

```{r}
df_wide <- df %>%
  pivot_wider(names_from  = year,
              values_from = c(co2, pm25))
```

### Calculate 2005-2010 change in CO2 and PM2.5 for each county

```{r}
df_wide <- df_wide %>%
  mutate(co2_change = co2_2010 - co2_2005,
         co2_rel_change = co2_change / co2_2005,
         co2_rel_change_perc = co2_rel_change * 100,
         pm25_change = pm25_2010 - pm25_2005,
         pm25_rel_change = pm25_change / pm25_2005,
         pm25_rel_change_perc = pm25_rel_change * 100)
```

### Create vector for counties with extreme values of `co2_rel_change_perc`

```{r}
# Filter for extreme counties
extreme_counties <- df_wide %>% 
  filter(abs(co2_rel_change_perc) >= 100)

# Get vector of corresponding fips codes
extreme_fips <- extreme_counties$county
```

### Filter `df_wide` and `co2_county` by removing extreme counties

```{r}
# Filter df_wide
df_wide <- df_wide %>% 
  filter(!(county %in% extreme_fips))

# Filter co2_county
co2_county <- co2_county %>% 
  filter(!(county %in% extreme_fips))
```

# Data visualization

### Aggregate CO2

```{r}
# Aggregate total CO2 for all counties by year
co2_total_year <- co2_county %>%
  group_by(year) %>%
  summarize(total_co2 = sum(co2, na.rm = TRUE))

# Check the result
head(co2_total_year)
```

### Plot aggregate CO2

```{r}
# Create graph of CO2 data
co2_plot <- co2_total_year %>%
  ggplot() +
  geom_line(aes(x = year, y = total_co2)) +
  geom_vline(aes(xintercept = 2005, linetype = "2005: Nonattainment Designation from NAAQS"), 
             color = "red") +
  labs(title = "Aggregated On-Road CO2 Emissions in the Contiguous U.S.",
       x = "Year",
       y = "CO2 (kilograms)",
       fill = "",
       linetype = "") +
  scale_linetype_manual(values = c("2005: Nonattainment Designation from NAAQS" = "dotted")) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Plot and save graph
co2_plot
ggsave("co2_plot.png",
       plot = co2_plot,
       bg = "white")
```

# Conclusion

### Save results

```{r}
# Save ordinary tabular datasets
write_parquet(df, "df.parquet")
write_parquet(co2_county, "co2_county.parquet")

# Save spatial dataset
st_write(counties_sf, "counties_sf.gpkg", delete_dsn = TRUE)
```