---
title: "04: Run plm"
author: "Ricky Truong"
date: "Fall 2026"
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
  html_document:
    df_print: paged
geometry: margin=1in
fontsize: 11pt
---

# Background

### Delete everything in environment

```{r, message = FALSE, warning = FALSE}
rm(list = ls())
```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(broom)
library(CausalArima)
library(colorspace)
library(ggnewscale)
library(ggpattern)
library(glue)
library(mgcv)
library(moderndive)
library(moments)
library(npreg)
library(plm)
library(plotly)
library(purrr)
library(scales)
library(scatterplot3d)
library(sf)
library(spdep)
library(splm)
library(tidyverse)
library(tigris)
library(tseries)
```

### Load data

```{r}
co2_county <- read_parquet("co2_county.parquet")
df <- read_parquet("df.parquet")
counties_sf <- st_read("counties_sf.gpkg")
```

# Analysis with `plm` library

### Prepare for `plm()` with `co2_county`

```{r}
# Rename variable to fips in co2_county
co2_county <- co2_county %>%
  rename(fips = county)

# Filter for years less than or equal to 2004 (to fit to pre-treatment period)
co2_county_2004 <- co2_county %>%
  filter(year <= 2004)

# Get shapefile of unique counties in co2_county_2004
fips_panel <- sort(unique(co2_county_2004$fips))
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  st_as_sf() %>%
  mutate(fips = str_pad(paste0(STATEFP, COUNTYFP), 5, "left", "0")) %>%
  filter(fips %in% fips_panel) %>%
  st_make_valid()

# Join co2_county_2004 with shapefile (to add polygons/geometry)
co2_county_2004 <- counties_sf %>%
  inner_join(co2_county_2004, by = c("fips" = "fips"))

# Create neighbor list using KNN to ensure each county has 4 neighbors
coords <- st_coordinates(st_centroid(counties_sf))
nb_knn <- knn2nb(knearneigh(coords, k = 4))

# Create listw object from nb_knn
listw_object <- nb2listw(nb_knn, style = "W")

# Ensure ordering matches plm()
regions <- counties_sf$fips
co2_county_2004 <- co2_county_2004 %>%
  filter(fips %in% regions) %>%
  mutate(fips = factor(fips, levels = regions)) %>%
  arrange(fips, year)

# Drop 1980 for every county (since there is no lag before)
df_one <- co2_county_2004 %>%
  arrange(fips, year) %>%
  group_by(fips) %>%
  mutate(
    log_co2 = log(co2),
    log_lag_co2 = log(lag(co2))
  ) %>%
  filter(!is.na(log_lag_co2)) %>%
  ungroup()
```

### Run `plm()` with `df_one`: log-log model with 1 year of lag (i.e., most simple model)

```{r}
# Create panel dataframe from df_one
pdf <- pdata.frame(df_one, index = c("fips", "year"))

# Fit model
model_1 <- plm(
  log_co2 ~ log_lag_co2,
  data = pdf,
  model = "within",
  effect = "individual"
)
summary(model_1)

# Evaluate model with AIC
log_lik_val <- as.numeric(model_1$logLik)
k <- length(coef(model_1))
AIC_val_1 <- -2 * log_lik_val + 2 * k
AIC_val_1
```

This model is $\log(Y_{i, t}) = \beta_1 \log(y_{i, t - 1}) + \mu_{i} + \varepsilon_{i, t}$, which implies $\mathbb{E}[\log(Y_{i, t}) \mid \text{past}] = \beta_1 \log(y_{i, t - 1}) + \mu_{i}$. We estimate $\widehat{\mathbb{E}}[\log(Y_{i, t}) \mid \text{past}] = \hat{\beta}_1 \log(y_{i, t - 1}) + \hat{\mu}_{i}$. We thus want $\hat{\beta}_1$ (which we can extract from the model) and $\hat{\mu}_i$ (which we can calculate for each $i$).

Let $y_{i} = \log(Y_{i, t})$ and $x_{i} = y_{i, t - 1}$. Our model is $y_{i} = \beta_1 x_{i} + \mu_{i} + \varepsilon_{i, t}$. Averaging over $t \in \{1, \ldots, T\}$, we have $\bar{y}_{i} = \beta_1 \bar{x}_{i} + \mu_{i} + \bar{\varepsilon}_{i, t}$. Since we assume $\mathbb{E}[\varepsilon_{i, t} \mid \text{past}] = 0$, for large $n$, $\bar{\varepsilon}_{i, t} \approx 0$, so $\bar{y}_{i} \approx \beta_1 \bar{x}_{i} + \mu_{i}$ $\implies$ $\mu_{i} \approx \bar{y}_{i} - \beta_1 \bar{x}_{i}$ $\implies$ $\hat{\mu}_{i} \approx \bar{y}_{i} - \hat{\beta}_{1} \bar{x}_{i}$.

### Use `model_1` to estimate counterfactual CO2 values

```{r}
# Drop geometry
df_one <- df_one %>%
  sf::st_drop_geometry()

# Get beta_1_hat
beta_hat <- coef(model_1)
beta_1_hat <- beta_hat["log_lag_co2"]

# Get mu_i_hat for each county
mu_hat_df <- df_one %>%
  group_by(fips) %>%
  summarize(
    mu_hat = mean(log_co2 - beta_1_hat * log_lag_co2, na.rm = TRUE),
    .groups = "drop"
  )

# Predict 2001 using observed 2000 CO2
pred_2001 <- df_one %>%
  filter(year == 2000) %>%
  left_join(mu_hat_df, by = "fips") %>%
  mutate(
    pred_year = 2001,
    pred_log_co2_2001 = beta_1_hat * log_co2 + mu_hat,
    pred_co2_2001 = exp(pred_log_co2_2001)
  ) %>%
  select(
    fips,
    pred_year,
    co2_2000 = co2,
    log_co2_2000 = log_co2,
    mu_hat,
    pred_log_co2_2001,
    pred_co2_2001
  )

# Display predicted vs. observed CO2 in 2001 for each county
df_one %>% 
  filter(year == 2001) %>% 
  rename(co2_2001 = co2) %>% 
  left_join(pred_2001, by = "fips") %>% 
  select(fips, co2_2000, pred_co2_2001, co2_2001)
```

### Run `plm()` with `df_three`: log-log model with 3 years of lag (i.e., more complicated model)

```{r}
# Drop 1980, 1981, 1982 for every county (since we're using three years of lag)
df_three <- co2_county_2004 %>%
  arrange(fips, year) %>%
  group_by(fips) %>%
  mutate(
    log_co2 = log(co2),
    log_lag_co2_one = lag(log_co2, 1),
    log_lag_co2_two = lag(log_co2, 2),
    log_lag_co2_three = lag(log_co2, 3)
  ) %>%
  filter(
    !is.na(log_lag_co2_one),
    !is.na(log_lag_co2_two),
    !is.na(log_lag_co2_three)
  ) %>%
  ungroup()

# Create panel dataframe from df_three
pdf <- pdata.frame(df_three, index = c("fips", "year"))

# Fit model via plm()
model_2 <- plm(
  log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
  data = pdf,
  model = "within",
  effect = "individual"
)
summary(model_2)
```

This model is $\log(Y_{i, t}) = \beta_1 \log(y_{i, t - 1}) + \beta_2 \log(y_{i, t - 2}) + \beta_3 \log(y_{i, t - 3}) + \mu_{i} + \varepsilon_{i, t}$, which implies $\mathbb{E}[\log(Y_{i, t}) \mid \text{past}] = \beta_1 \log(y_{i, t - 1}) + \beta_2 \log(y_{i, t - 2}) + \beta_3 \log(y_{i, t - 3}) + \mu_{i}$. We estimate $\widehat{\mathbb{E}}[\log(Y_{i, t}) \mid \text{past}] = \hat{\beta}_1 \log(y_{i, t - 1}) + \hat{\beta}_2 \log(y_{i, t - 2}) + \hat{\beta}_3 \log(y_{i, t - 3}) + \hat{\mu}_{i}$. We thus want $\hat{\beta}_1$, $\hat{\beta}_2$, and $\hat{\beta}_3$ (which we can extract from the model) and $\hat{\mu}_i$ (which we can calculate for each $i$).

Let $y_{i, t} = \log(Y_{i, t})$ and $x_{i, t - j} = y_{i, t - j}$. Our model is $y_{i, t} = \beta_1 x_{i, t - 1} + \beta_2 x_{i, t - 2} + \beta_3 x_{i, t - 3} + \mu_{i} + \varepsilon_{i, t}$. Averaging over $t \in \{1, \ldots, T\}$, we have $\bar{y}_{i} = \beta_1 \bar{x}_{i - 1} + \beta_2 \bar{x}_{i - 2} + \beta_3 \bar{x}_{i - 3} + \mu_{i} + \bar{\varepsilon}_{i, t}$. Since we assume $\mathbb{E}[\varepsilon_{i, t} \mid \text{past}] = 0$, for large $n$, $\bar{\varepsilon}_{i, t} \approx 0$, so $\bar{y}_{i} \approx \beta_1 \bar{x}_{i - 1} + \beta_2 \bar{x}_{i - 2} + \beta_3 \bar{x}_{i - 3} + \mu_{i}$ $\implies$ $\mu_{i} \approx \bar{y}_{i} - \beta_1 \bar{x}_{i - 1} - \beta_2 \bar{x}_{i - 2} - \beta_3 \bar{x}_{i - 3}$ $\implies$ $\hat{\mu}_{i} \approx \bar{y}_{i} - \hat{\beta}_{1} \bar{x}_{i - 1} - \hat{\beta}_{2} \bar{x}_{i - 2} - \hat{\beta}_{3} \bar{x}_{i - 3}$.

### Use `model_2` to estimate counterfactual CO2 values

```{r}
# Drop geometry
df_clean <- df_three %>%
  sf::st_drop_geometry()

# Get beta estimates
beta_hat <- coef(model_2)
beta_1_hat <- beta_hat["log_lag_co2_one"]
beta_2_hat <- beta_hat["log_lag_co2_two"]
beta_3_hat <- beta_hat["log_lag_co2_three"]

# Get mu_i_hat for each county
mu_hat_df <- df_clean %>%
  group_by(fips) %>%
  summarize(
    mu_hat = mean(
      log_co2 -
        beta_1_hat * log_lag_co2_one -
        beta_2_hat * log_lag_co2_two -
        beta_3_hat * log_lag_co2_three,
      na.rm = TRUE
    ),
    .groups = "drop"
  )

# Prepare full county-year data without geometry
df_all <- co2_county %>%
  sf::st_drop_geometry() %>%
  mutate(
    fips = as.character(fips),
    log_co2 = log(co2)
  ) %>%
  arrange(fips, year)

# Create wide data with observed log CO2 values
pred_wide <- df_all %>%
  filter(year >= 2002, year <= 2010) %>%
  select(fips, year, log_co2) %>%
  tidyr::pivot_wider(
    names_from = year,
    values_from = log_co2,
    names_prefix = "log_co2_"
  ) %>%
  left_join(mu_hat_df, by = "fips")

# For each year in 2005-2010, recursively predict log CO2
for (yr in 2005:2010) {
  
  # Define lag column names
  lag_1 <- paste0("log_co2_", yr - 1)
  lag_2 <- paste0("log_co2_", yr - 2)
  lag_3 <- paste0("log_co2_", yr - 3)
  pred_col <- paste0("pred_log_co2_", yr)
  
  # Predict log CO2 for current year
  pred_wide <- pred_wide %>%
    mutate(
      !!pred_col :=
        beta_1_hat * .data[[lag_1]] +
        beta_2_hat * .data[[lag_2]] +
        beta_3_hat * .data[[lag_3]] +
        mu_hat
    )
  
  # Replace log_co2_yr with predicted value so future years use recursive predictions
  pred_wide <- pred_wide %>%
    mutate(
      !!paste0("log_co2_", yr) := .data[[pred_col]]
    )
}

# Convert predictions to long format
pred_2005_2010 <- pred_wide %>%
  select(fips, starts_with("pred_log_co2_")) %>%
  tidyr::pivot_longer(
    cols = starts_with("pred_log_co2_"),
    names_to = "pred_year",
    values_to = "pred_log_co2"
  ) %>%
  mutate(
    pred_year = as.numeric(stringr::str_remove(pred_year, "pred_log_co2_")),
    pred_co2 = exp(pred_log_co2)
  )

# Join observed CO2 values for 2005-2010
pred_vs_obs <- df_all %>%
  filter(year >= 2005, year <= 2010) %>%
  select(fips, year, observed_co2 = co2, observed_log_co2 = log_co2) %>%
  left_join(
    pred_2005_2010,
    by = c("fips" = "fips", "year" = "pred_year")
  )

# Display predicted vs. observed CO2 in 2005-2010 for each county
pred_vs_obs %>%
  select(fips, year, observed_co2, pred_co2)
```

### Plot predicted vs. observed CO2 for individual county and for total (from `model_2`)

```{r}
# Plot for individual county
pred_vs_obs %>% 
  filter(fips == "01053") %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = observed_co2, color = "Observed CO2")) +
  geom_line(aes(y = pred_co2, color = "Predicted CO2")) +
  scale_color_manual(values = c("Observed CO2" = "blue", "Predicted CO2" = "red"))

# Plot for total/aggregated
pred_vs_obs %>% 
  group_by(year) %>% 
  summarize(total_observed_co2 = sum(observed_co2, na.rm = TRUE),
            total_pred_co2 = sum(pred_co2, na.rm = TRUE)) %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = total_observed_co2, color = "Total Observed CO2")) +
  geom_line(aes(y = total_pred_co2, color = "Total Predicted CO2")) +
  scale_color_manual(values = c("Total Observed CO2" = "blue", "Total Predicted CO2" = "red"))
```

### Analyze Normality of residuals (from `model_2`) for one county 

```{r}
# Define function to get residuals for a county
specified_fips <- "01053"
get_county_residuals_three_lag <- function(
    fips_use = specified_fips,
    co2_train = co2_county_2004
) {
  
  # Build 3-lag training data (pre-treatment only, up to 2004)
  df_three <- co2_train %>%
    sf::st_drop_geometry() %>%
    mutate(
      fips = as.character(fips),
      log_co2 = log(co2)
    ) %>%
    arrange(fips, year) %>%
    group_by(fips) %>%
    mutate(
      log_lag_co2_one = lag(log_co2, 1),
      log_lag_co2_two = lag(log_co2, 2),
      log_lag_co2_three = lag(log_co2, 3)
    ) %>%
    filter(
      !is.na(log_lag_co2_one),
      !is.na(log_lag_co2_two),
      !is.na(log_lag_co2_three)
    ) %>%
    ungroup()
  
  # Convert to panel dataframe
  pdf <- pdata.frame(df_three, index = c("fips", "year"))
  
  # Fit within / county fixed effects model
  model_2 <- plm(
    log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
    data = pdf,
    model = "within",
    effect = "individual"
  )
  
  # Extract beta estimates
  beta_hat <- coef(model_2)
  beta_1_hat <- beta_hat["log_lag_co2_one"]
  beta_2_hat <- beta_hat["log_lag_co2_two"]
  beta_3_hat <- beta_hat["log_lag_co2_three"]
  
  # Reconstruct county fixed effects mu_i
  mu_hat_df <- df_three %>%
    group_by(fips) %>%
    summarize(
      mu_hat = mean(
        log_co2 -
          beta_1_hat * log_lag_co2_one -
          beta_2_hat * log_lag_co2_two -
          beta_3_hat * log_lag_co2_three,
        na.rm = TRUE
      ),
      .groups = "drop"
    )
  
  # Get fitted values and residuals for county
  county_resid_df <- df_three %>%
    filter(fips == fips_use) %>%
    left_join(mu_hat_df, by = "fips") %>%
    mutate(
      fitted_log_co2 =
        beta_1_hat * log_lag_co2_one +
        beta_2_hat * log_lag_co2_two +
        beta_3_hat * log_lag_co2_three +
        mu_hat,
      
      residual_log_co2 = log_co2 - fitted_log_co2,
      
      fitted_co2 = exp(fitted_log_co2),
      residual_co2 = co2 - fitted_co2
    ) %>%
    select(
      fips, year, co2, log_co2,
      fitted_log_co2, residual_log_co2,
      fitted_co2, residual_co2,
      log_lag_co2_one, log_lag_co2_two, log_lag_co2_three,
      mu_hat
    )
  
  return(list(
    model = model_2,
    beta_hat = beta_hat,
    mu_hat_df = mu_hat_df,
    county_residuals = county_resid_df
  ))
}

# Get residuals for fips 01053
res_01053 <- get_county_residuals_three_lag(fips_use = "01053")

# Summarize residuals
res_01053$county_residuals
summary(res_01053$model)

# Plot histogram of residuals
resid_01053 <- res_01053$county_residuals$residual_log_co2
hist(resid_01053, breaks = 10, main = "Residuals for FIPS 01053",
     xlab = "Log-scale residuals")

# Plot Q-Q plot
qqnorm(resid_01053)
qqline(resid_01053)

# Run Shapiro-Wilk test (high p indicates Normality)
shapiro.test(resid_01053)

# Plot observed and fitted values
res_01053$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"), linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"), size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 01053",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

### Analyze Normality of residuals (from `model_2`) for all counties

```{r}
# Run on all counties
county_diag_df <- df_three %>%
  group_split(fips) %>%
  map_dfr(function(df_county) {
    
    # Get county residuals
    county_res <- df_county %>%
      left_join(mu_hat_df, by = "fips") %>%
      mutate(
        fitted =
          beta_1_hat * log_lag_co2_one +
          beta_2_hat * log_lag_co2_two +
          beta_3_hat * log_lag_co2_three +
          mu_hat,
        
        resid = log_co2 - fitted
      )
    
    r <- county_res$resid
    
    tibble(
      fips = unique(df_county$fips),
      n = length(r),
      mean_resid = mean(r),
      sd_resid = sd(r),
      skewness = moments::skewness(r),
      kurtosis = moments::kurtosis(r),
      shapiro_p = shapiro.test(r)$p.value,
      jb_p = jarque.bera.test(r)$p.value,
      acf1 = acf(r, plot = FALSE)$acf[2]
    )
  })

# Plot histogram of p-values from Shapiro-Wilk test
ggplot(county_diag_df, aes(x = shapiro_p)) +
  geom_histogram(bins = 30) +
  theme_minimal()

# Get proportion of counties that have low p-value (non-Normality)
mean(county_diag_df$shapiro_p < 0.01)

# Display counties with low p-value (non-Normality)
county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1:10)
```

### Analyze Normality of residuals (from `model_2`) for a "bad" county

```{r}
# Get fips and residuals of a county with low p-value (non-Normality)
bad_fips <- county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1) %>%
  pull(fips)
bad_res <- get_county_residuals_three_lag(
  fips_use = bad_fips
)$county_residuals

# Plot residuals over time
ggplot(bad_res, aes(x = year, y = residual_log_co2)) +
  geom_line() +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = paste("Residual Time Series:", bad_fips),
    x = "Year",
    y = "Residual"
  ) +
  theme_minimal()

# Plot histogram of residuals
ggplot(bad_res, aes(x = residual_log_co2)) +
  geom_histogram(bins = 10) +
  labs(
    title = paste("Residual Histogram:", bad_fips),
    x = "Residual"
  ) +
  theme_minimal()

# Plot Q-Q plot
ggplot(bad_res, aes(sample = residual_log_co2)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = paste("QQ Plot:", bad_fips)
  ) +
  theme_minimal()

# Plot observed and fitted values
res_29083 <- get_county_residuals_three_lag(fips_use = "29083")
resid_29083 <- res_29083$county_residuals$residual_log_co2
res_29083$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"),
            linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"),
             size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 29083",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

# Conclusion

### Save results

```{r}
write_csv(df_one, "df_one.csv")
write_csv(df_three, "df_three.csv")
```