---
title: "05: Run splm"
author: "Ricky Truong"
date: "Fall 2026"
output:
  pdf_document:
    fig_width: 5
    fig_height: 4
  html_document:
    df_print: paged
geometry: margin=1in
fontsize: 11pt
---

# Background

### Delete everything in environment

```{r, message = FALSE, warning = FALSE}
rm(list = ls())
```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(broom)
library(CausalArima)
library(colorspace)
library(ggnewscale)
library(ggpattern)
library(glue)
library(mgcv)
library(moderndive)
library(moments)
library(npreg)
library(plm)
library(plotly)
library(purrr)
library(scales)
library(scatterplot3d)
library(sf)
library(spdep)
library(splm)
library(tidyverse)
library(tigris)
library(tseries)
```

### Load data

```{r}
co2_county <- read_parquet("co2_county.parquet")
df <- read_parquet("df.parquet")
counties_sf <- st_read("counties_sf.gpkg")
df_one <- read_csv("df_one.csv")
df_three <- read_csv("df_three.csv")
```

# Analysis with `splm` library

### Prepare for `spml()` with `co2_county`

```{r}
# Rename variable to fips in co2_county
co2_county <- co2_county %>%
  rename(fips = county)

# Filter for years less than or equal to 2004 (to fit to pre-treatment period)
co2_county_2004 <- co2_county %>%
  filter(year <= 2004)

# Get shapefile of unique counties in co2_county_2004
fips_panel <- sort(unique(co2_county_2004$fips))
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  st_as_sf() %>%
  mutate(fips = str_pad(paste0(STATEFP, COUNTYFP), 5, "left", "0")) %>%
  filter(fips %in% fips_panel) %>%
  st_make_valid()

# Join co2_county_2004 with shapefile (to add polygons/geometry)
co2_county_2004 <- counties_sf %>%
  inner_join(co2_county_2004, by = c("fips" = "fips"))

# Create neighbor list using KNN to ensure each county has 4 neighbors
coords <- st_coordinates(st_centroid(counties_sf))
nb_knn <- knn2nb(knearneigh(coords, k = 4))

# Create listw object from nb_knn
listw_object <- nb2listw(nb_knn, style = "W")
```

### Recreate `model_2` with `spml()` (i.e., three years of lag and spatial errors)

```{r}
# Create panel dataframe from df_three
pdf <- pdata.frame(df_three, index = c("fips", "year"))

# Fit model via spml()
model_3 <- spml(
  log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
  data = pdf,
  listw = listw_object,
  model = "within",
  effect = "individual",
  lag = FALSE
)
summary(model_3)
```

This model is $\log(Y_{i, t}) = \alpha + \beta_1 \log(y_{i, t - 1}) + \beta_2 \log(y_{i, t - 2}) + \beta_3 \log(y_{i, t - 3}) + \mu_{i} + u_{i, t}$, where $u_t = \rho W u_t + \varepsilon_t$ (equivalently, $(I - \rho W) u_t = \varepsilon_t$, which implies $u_t = (I - \rho W)^{-1} \varepsilon_t$). 

Let $y_{i, t} = \log(Y_{i, t})$ and $x_{i, t - j} = y_{i, t - j}$. Our model is $y_{i, t} = \alpha + \beta_1 x_{i, t - 1} + \beta_2 x_{i, t - 2} + \beta_3 x_{i, t - 3} + \mu_{i} + u_{i, t}$, and we assume $\mathbb{E}[y_{i, t} \mid \text{past}] = \alpha + \beta_1 x_{i, t - 1} + \beta_2 x_{i, t - 2} + \beta_3 x_{i, t - 3} + \mu_{i}$, so we estimate $\widehat{\mathbb{E}}[\log(Y_{i, t}) \mid \text{past}] = \hat{\alpha} + \hat{\beta}_1 \log(y_{i, t - 1}) + \hat{\beta}_2 \log(y_{i, t - 2}) + \hat{\beta}_3 \log(y_{i, t - 3}) + \hat{\mu}_{i}$.

We thus want $\hat{\alpha}$, $\hat{\beta}_1$, $\hat{\beta}_2$, $\hat{\beta}_3$, and $\hat{\mu}_i$ (which we can extract from the model). Specifically, for an `spml()` object, we can run `effects()` and absorb $\hat{\alpha}$ into $\hat{\mu}_i$.

### Get mu_i_hat for each county

```{r}
# Extract fixed effects object from spml model
effects_3 <- effects(model_3)

# Extract common intercept
intercept_hat <- effects_3$INTTable["(Intercept)", "Estimate"]

# Extract county-specific fixed effect deviations
mu_hat_raw <- effects_3$SETable[, "Estimate"]

# Get county ordering from the pdata.frame used to fit model_3
fips_order <- unique(as.character(index(pdf)$fips))

# Build county-level df for mu_hat
mu_hat_df <- tibble(
  fips = fips_order,
  mu_hat = as.numeric(intercept_hat + mu_hat_raw)
)
```

### Use `model_3` to estimate counterfactual CO2 values

```{r}
# Drop geometry
df_clean <- df_three %>%
  sf::st_drop_geometry()

# Get beta estimates
beta_hat <- coef(model_3)
beta_1_hat <- beta_hat["log_lag_co2_one"]
beta_2_hat <- beta_hat["log_lag_co2_two"]
beta_3_hat <- beta_hat["log_lag_co2_three"]

# Prepare full county-year data without geometry
df_all <- co2_county %>%
  sf::st_drop_geometry() %>%
  mutate(
    fips = as.character(fips),
    log_co2 = log(co2)
  ) %>%
  arrange(fips, year)

# Create wide data with observed log CO2 values
pred_wide <- df_all %>%
  filter(year >= 2002, year <= 2010) %>%
  select(fips, year, log_co2) %>%
  tidyr::pivot_wider(
    names_from = year,
    values_from = log_co2,
    names_prefix = "log_co2_"
  ) %>%
  left_join(mu_hat_df, by = "fips")

# For each year in 2005-2010, recursively predict log CO2
for (yr in 2005:2010) {
  
  # Define lag column names
  lag_1 <- paste0("log_co2_", yr - 1)
  lag_2 <- paste0("log_co2_", yr - 2)
  lag_3 <- paste0("log_co2_", yr - 3)
  pred_col <- paste0("pred_log_co2_", yr)
  
  # Predict log CO2 for current year
  pred_wide <- pred_wide %>%
    mutate(
      !!pred_col :=
        beta_1_hat * .data[[lag_1]] +
        beta_2_hat * .data[[lag_2]] +
        beta_3_hat * .data[[lag_3]] +
        mu_hat
    )
  
  # Replace log_co2_yr with predicted value so future years use recursive predictions
  pred_wide <- pred_wide %>%
    mutate(
      !!paste0("log_co2_", yr) := .data[[pred_col]]
    )
}

# Convert predictions to long format
pred_2005_2010 <- pred_wide %>%
  select(fips, starts_with("pred_log_co2_")) %>%
  tidyr::pivot_longer(
    cols = starts_with("pred_log_co2_"),
    names_to = "pred_year",
    values_to = "pred_log_co2"
  ) %>%
  mutate(
    pred_year = as.numeric(stringr::str_remove(pred_year, "pred_log_co2_")),
    pred_co2 = exp(pred_log_co2)
  )

# Join observed CO2 values for 2005-2010
pred_vs_obs <- df_all %>%
  filter(year >= 2005, year <= 2010) %>%
  select(fips, year, observed_co2 = co2, observed_log_co2 = log_co2) %>%
  left_join(
    pred_2005_2010,
    by = c("fips" = "fips", "year" = "pred_year")
  )

# Display predicted vs. observed CO2 in 2005-2010 for each county
pred_vs_obs %>%
  select(fips, year, observed_co2, pred_co2)
```

### Plot predicted vs. observed CO2 for individual county and for total (from `model_3`)

```{r}
# Plot for individual county
pred_vs_obs %>% 
  filter(fips == "01053") %>% 
  ggplot(aes(x = year)) + 
  geom_line(aes(y = observed_co2, color = "Observed CO2")) +
  geom_line(aes(y = pred_co2, color = "Predicted CO2")) +
  scale_color_manual(values = c("Observed CO2" = "blue", "Predicted CO2" = "red"))

# Plot for total/aggregated
pred_vs_obs %>% 
  group_by(year) %>% 
  summarize(total_observed_co2 = sum(observed_co2, na.rm = TRUE),
            total_pred_co2 = sum(pred_co2, na.rm = TRUE)) %>% 
  ggplot(aes(x = year)) +
  geom_line(aes(y = total_observed_co2, color = "Total Observed CO2")) +
  geom_line(aes(y = total_pred_co2, color = "Total Predicted CO2")) +
  scale_color_manual(values = c("Total Observed CO2" = "blue", "Total Predicted CO2" = "red"))
```

### Analyze Normality of residuals (from `model_3`) for one county 

```{r}
# Define function to get residuals for a county
specified_fips <- "01053"
get_county_residuals_three_lag <- function(
    fips_use = specified_fips,
    co2_train = co2_county_2004
) {
  
  # Build 3-lag training data (pre-treatment only, up to 2004)
  df_three <- co2_train %>%
    sf::st_drop_geometry() %>%
    mutate(
      fips = as.character(fips),
      log_co2 = log(co2)
    ) %>%
    arrange(fips, year) %>%
    group_by(fips) %>%
    mutate(
      log_lag_co2_one = lag(log_co2, 1),
      log_lag_co2_two = lag(log_co2, 2),
      log_lag_co2_three = lag(log_co2, 3)
    ) %>%
    filter(
      !is.na(log_lag_co2_one),
      !is.na(log_lag_co2_two),
      !is.na(log_lag_co2_three)
    ) %>%
    ungroup()
  
  # Convert to panel dataframe
  pdf <- pdata.frame(df_three, index = c("fips", "year"))
  
  # Fit within / county fixed effects model
  model_3 <- spml(
    log_co2 ~ log_lag_co2_one + log_lag_co2_two + log_lag_co2_three,
    data = pdf,
    listw = listw_object,
    model = "within",
    effect = "individual",
    lag = FALSE
  )
  
  # Extract beta estimates
  beta_hat <- coef(model_3)
  beta_1_hat <- beta_hat["log_lag_co2_one"]
  beta_2_hat <- beta_hat["log_lag_co2_two"]
  beta_3_hat <- beta_hat["log_lag_co2_three"]
  
  # Reconstruct county fixed effects mu_i
  effects_3 <- effects(model_3)
  intercept_hat <- effects_3$INTTable["(Intercept)", "Estimate"]
  mu_hat_raw <- effects_3$SETable[, "Estimate"]
  fips_order <- unique(as.character(index(pdf)$fips))
  mu_hat_df <- tibble(
    fips = fips_order,
    mu_hat = as.numeric(intercept_hat + mu_hat_raw)
  )
  
  # Get fitted values and residuals for county
  county_resid_df <- df_three %>%
    filter(fips == fips_use) %>%
    left_join(mu_hat_df, by = "fips") %>%
    mutate(
      fitted_log_co2 =
        beta_1_hat * log_lag_co2_one +
        beta_2_hat * log_lag_co2_two +
        beta_3_hat * log_lag_co2_three +
        mu_hat,
      
      residual_log_co2 = log_co2 - fitted_log_co2,
      
      fitted_co2 = exp(fitted_log_co2),
      residual_co2 = co2 - fitted_co2
    ) %>%
    select(
      fips, year, co2, log_co2,
      fitted_log_co2, residual_log_co2,
      fitted_co2, residual_co2,
      log_lag_co2_one, log_lag_co2_two, log_lag_co2_three,
      mu_hat
    )
  
  return(list(
    model = model_3,
    beta_hat = beta_hat,
    mu_hat_df = mu_hat_df,
    county_residuals = county_resid_df
  ))
}

# Get residuals for fips 01053
res_01053 <- get_county_residuals_three_lag(fips_use = "01053")

# Summarize residuals
res_01053$county_residuals
summary(res_01053$model)

# Plot histogram of residuals
resid_01053 <- res_01053$county_residuals$residual_log_co2
hist(resid_01053, breaks = 10, main = "Residuals for FIPS 01053",
     xlab = "Log-scale residuals")

# Plot Q-Q plot
qqnorm(resid_01053)
qqline(resid_01053)

# Run Shapiro-Wilk test (high p indicates Normality)
shapiro.test(resid_01053)

# Plot observed and fitted values
res_01053$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"), linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"), size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 01053",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

### Analyze Normality of residuals (from `model_3`) for all counties

```{r}
# Run on all counties
county_diag_df <- df_three %>%
  group_split(fips) %>%
  map_dfr(function(df_county) {
    
    # Get county residuals
    county_res <- df_county %>%
      left_join(mu_hat_df, by = "fips") %>%
      mutate(
        fitted =
          beta_1_hat * log_lag_co2_one +
          beta_2_hat * log_lag_co2_two +
          beta_3_hat * log_lag_co2_three +
          mu_hat,
        
        resid = log_co2 - fitted
      )
    
    r <- county_res$resid
    
    tibble(
      fips = unique(df_county$fips),
      n = length(r),
      mean_resid = mean(r),
      sd_resid = sd(r),
      skewness = moments::skewness(r),
      kurtosis = moments::kurtosis(r),
      shapiro_p = shapiro.test(r)$p.value,
      jb_p = jarque.bera.test(r)$p.value,
      acf1 = acf(r, plot = FALSE)$acf[2]
    )
  })

# Plot histogram of p-values from Shapiro-Wilk test
ggplot(county_diag_df, aes(x = shapiro_p)) +
  geom_histogram(bins = 30) +
  theme_minimal()

# Get proportion of counties that have low p-value (non-Normality)
mean(county_diag_df$shapiro_p < 0.01)

# Display counties with low p-value (non-Normality)
county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1:10)
```

### Analyze Normality of residuals (from `model_3`) for a "bad" county

```{r}
# Get fips and residuals of a county with low p-value (non-Normality)
bad_fips <- county_diag_df %>%
  arrange(shapiro_p) %>%
  slice(1) %>%
  pull(fips)
bad_res <- get_county_residuals_three_lag(
  fips_use = bad_fips
)$county_residuals

# Plot residuals over time
ggplot(bad_res, aes(x = year, y = residual_log_co2)) +
  geom_line() +
  geom_point(size = 2) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(
    title = paste("Residual Time Series:", bad_fips),
    x = "Year",
    y = "Residual"
  ) +
  theme_minimal()

# Plot histogram of residuals
ggplot(bad_res, aes(x = residual_log_co2)) +
  geom_histogram(bins = 10) +
  labs(
    title = paste("Residual Histogram:", bad_fips),
    x = "Residual"
  ) +
  theme_minimal()

# Plot Q-Q plot
ggplot(bad_res, aes(sample = residual_log_co2)) +
  stat_qq() +
  stat_qq_line() +
  labs(
    title = paste("QQ Plot:", bad_fips)
  ) +
  theme_minimal()

# Plot observed and fitted values
res_29083 <- get_county_residuals_three_lag(fips_use = "29083")
resid_29083 <- res_29083$county_residuals$residual_log_co2
res_29083$county_residuals %>%
  ggplot(aes(x = year)) +
  geom_line(aes(y = fitted_log_co2, color = "Fitted log(CO2)"),
            linewidth = 1) +
  geom_point(aes(y = log_co2, color = "Observed log(CO2)"),
             size = 2) +
  labs(
    title = "Observed vs. Fitted log(CO2) for FIPS 29083",
    x = "Year",
    y = "log(CO2)",
    color = ""
  ) +
  theme_minimal()
```

```{r}
# Extract fitted values
model_3$fitted.values
```

```{r}
# Built-in fitted values from spml
fitted_builtin_df <- tibble(
  fips_year = names(model_3$fitted.values),
  fitted_builtin = as.numeric(model_3$fitted.values)
) %>%
  tidyr::separate(
    fips_year,
    into = c("fips", "year"),
    sep = "-",
    convert = TRUE
  ) %>%
  mutate(fips = sprintf("%05d", fips))

fitted_manual_df <- df_three %>%
  sf::st_drop_geometry() %>%
  mutate(fips = as.character(fips)) %>%
  left_join(mu_hat_df, by = "fips") %>%
  mutate(
    fitted_manual =
      beta_1_hat * log_lag_co2_one +
      beta_2_hat * log_lag_co2_two +
      beta_3_hat * log_lag_co2_three +
      mu_hat
  ) %>%
  select(fips, year, log_co2, fitted_manual)

check_01053 <- fitted_manual_df %>%
  filter(fips == "01053") %>%
  left_join(fitted_builtin_df, by = c("fips", "year")) %>%
  mutate(diff = fitted_manual - fitted_builtin)

check_01053
summary(check_01053$diff)
max(abs(check_01053$diff), na.rm = TRUE)
```

```{r}
check_all <- fitted_manual_df %>%
  left_join(fitted_builtin_df, by = c("fips", "year")) %>%
  mutate(diff = fitted_manual - fitted_builtin)

summary(check_all$diff)
max(abs(check_all$diff), na.rm = TRUE)
```

```{r}
effects_3 <- effects(model_3)

head(rownames(effects_3$SETable))
head(unique(as.character(index(pdf)$fips)))
```

```{r}
fips_from_fitted <- names(model_3$fitted.values) %>%
  stringr::str_extract("^[0-9]{5}") %>%
  unique()
```

```{r}
effects(model_3)
```