---
title: "Aggregation"
author: "Ricky Truong"
date: "August 2025"

fontsize: 11pt
geometry: margin=1in

output:
  pdf_document:
    fig_width: 5
    fig_height: 4
---

# Background

### Delete everything in environment

# ```{r, message = FALSE, warning = FALSE}
# rm(list = ls())
# ```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(sf)
library(tidyverse)
```

### Load data for crosswalk (between block group and ZCTA)

```{r}
crosswalk <- read_parquet("nhgis__block2zcta__2010.parquet")
```

### Load data for PM2.5 (at ZCTA level)

```{r}
# Create path to the .parquet files
pm25_path <- "dataverse_files"

# List all .parquet files for years 2000–2020
file_list <- list.files(path = pm25_path,
                        pattern = "^pm25__ushap__zcta_yearly__20\\d{2}\\.parquet$",
                        full.names = TRUE)

# Read and combine all files into data frame
pm25 <- file_list %>%
  lapply(read_parquet) %>%
  bind_rows()
```

### Load data for CO2 (at census-block-group level)

```{r}
# Create path to the .gdb directory
co2_path <- "CMS_DARTE_V2_1735/data/DARTE_v2.gdb"

# List the available layers (feature classes) inside the .gdb
st_layers(dsn = co2_path)

# Load the specific layer
co2 <- st_read(dsn = co2_path, layer = "DARTE_v2_blockgroup_kgco2_1980_2017")
```

# Data wrangling for `crosswalk`

### Add separate variables for `block` and `block_group`

```{r}
crosswalk <- crosswalk %>%
  rename(fips_code = block) %>%
  mutate(block_group = substring(fips_code, 12, 12),
         block = substring(fips_code, 13, 15))
```

### Calculate weights

```{r}
crosswalk <- crosswalk %>%
  group_by(state, county, tract, block_group) %>%
  mutate(block_group_population = sum(total_pop),
         block_fraction = total_pop / block_group_population)
```

### Add separate variable for `fips_code_no_block` (to combine with `co2`)

```{r}
crosswalk <- crosswalk %>%
  mutate(fips_code_no_block = substring(fips_code, 1, 12))
```

# Data wrangling for `co2`

### Convert `co2` to tidy format

```{r}
co2 <- co2 %>%
  pivot_longer(cols = starts_with("kgco2_"),
               names_to = "year",
               names_prefix = "kgco2_",
               values_to = "value") %>%
  mutate(year = as.integer(year))
```

# Previewing

### Preview `crosswalk`

```{r}
head(crosswalk)
```

### Preview `pm25`

```{r}
head(pm25)
```

### Preview `co2`

```{r}
head(co2)
```

# Apply `crosswalk` 

### Search `crosswalk` for value in `pm25` as a sanity check

```{r}
crosswalk %>%
  filter(zcta == "01040")
```

### Search `crosswalk` for value in `co2` as a sanity check

```{r}
crosswalk %>%
  filter(fips_code_no_block == "010810416001")
```

### Combine `crosswalk` and `co2`

```{r}
# Select for only necessary variables to avoid large data sets
crosswalk <- crosswalk %>%
  ungroup() %>%
  select(fips_code, zcta, block_fraction, fips_code_no_block)

co2 <- co2 %>%
  ungroup() %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  select(GEOID, year, value)

# Combine via inner_join()
df <- inner_join(crosswalk, co2, join_by("fips_code_no_block" == "GEOID"))

# Wrangle df
df <- df %>%
  rename(block_group_value = value) %>%
  mutate(block_value = block_group_value * block_fraction) %>%
  select(-c(fips_code_no_block, block_fraction))

# Preview df
head(df, 100)
```

### Double check output (again, as a sanity check)

```{r}
# For block group of 010010201001000 in 1980, the value should be 2514107
co2 %>%
  filter(GEOID == "010010201001", year == 1980)
```

# Applying `pm25`

### Combine `pm25` and `df`

```{r}
# Rename block_value to CO2 and aggregate from block-year to ZCTA-year
df_zcta_year <- df %>%
  rename(CO2 = block_value) %>%
  group_by(zcta, year) %>%
  summarise(CO2 = sum(CO2, na.rm = TRUE), .groups = "drop")

# Combine via inner_join()
aggregated <- inner_join(pm25, df_zcta_year, join_by("zcta", "year"))
```

# Double check extreme values

### Preview `aggregated`

```{r}
# Preview aggregated with no organization
head(aggregated, 100)

# Preview aggregated with organization
aggregated %>%
  arrange(zcta) %>%
  select(year, zcta, PM25, CO2) %>%
  head(100)

# Filter for rows with missing values for PM25
missing_pm25 <- aggregated %>%
  filter(is.na(PM25))

missing_pm25

# Filter for rows with missing values for CO2
missing_co2 <- aggregated %>%
  filter(is.na(CO2))

missing_co2

# Filter for rows with zero for PM25
zero_pm25 <- aggregated %>%
  filter(PM25 == 0)

zero_pm25

# Filter for rows with zero for CO2
zero_co2 <- aggregated %>%
  filter(CO2 == 0)

zero_co2
```

### Search `pm25` for missing value in `aggregated` as a sanity check

```{r}
# For ZCTA of 00601 in 2000, the value should be NA
pm25 %>%
  filter(zcta == "00601", year == 2000)
```

### Search `df` for zero value in `aggregated` as a sanity check

```{r}
# For ZCTA of 00601 in 2000, the values should be 0
df %>%
  filter(zcta == "00601", year == 2000)
```

# Exporting files

### Export `aggregated`

```{r}
write.csv(aggregated, "~/aggregated.csv", row.names = FALSE)
```