---
title: "Draft"
author: "Ricky Truong"
date: "December 2025"

fontsize: 11pt
geometry: margin=1in

output:
  pdf_document:
    fig_width: 5
    fig_height: 4
---

# Background

### Delete everything in environment

```{r, message = FALSE, warning = FALSE}
rm(list = ls())
```

### Load libraries

```{r, message = FALSE, warning = FALSE}
library(arrow)
library(broom)
library(CausalArima)
library(colorspace)
library(glue)
library(mgcv)
library(moderndive)
library(scales)
library(scatterplot3d)
library(sf)
library(tidyverse)
library(tigris)
```

### Load data for county shapefile (from `library(tigris)`)

```{r}
# Load county shapefile (includes FIPS codes and geometry)
counties_sf <- counties(cb = TRUE, resolution = "20m", year = 2020) %>%
  mutate(fips = paste0(STATEFP, COUNTYFP))

# Filter for contiguous U.S.
contiguous_fips <- setdiff(sprintf("%02d", 1:56),
                           c("02", "15", "60", "66", "69", "72", "78"))
counties_sf <- counties_sf %>%
  filter(STATEFP %in% contiguous_fips)
```

### Load data for crosswalk (between ZCTA and county)

```{r}
xwalk22 <- read_parquet("us_xwalks__census__zcta2county_yearly__2022.parquet")
xwalk21 <- read_parquet("us_xwalks__census__zcta2county_yearly__2021.parquet")
xwalk20 <- read_parquet("us_xwalks__census__zcta2county_yearly__2020.parquet")
xwalk19 <- read_parquet("us_xwalks__census__zcta2county_yearly__2019.parquet")
xwalk18 <- read_parquet("us_xwalks__census__zcta2county_yearly__2018.parquet")
xwalk17 <- read_parquet("us_xwalks__census__zcta2county_yearly__2017.parquet")
xwalk16 <- read_parquet("us_xwalks__census__zcta2county_yearly__2016.parquet")
xwalk15 <- read_parquet("us_xwalks__census__zcta2county_yearly__2015.parquet")
xwalk14 <- read_parquet("us_xwalks__census__zcta2county_yearly__2014.parquet")
xwalk13 <- read_parquet("us_xwalks__census__zcta2county_yearly__2013.parquet")
xwalk12 <- read_parquet("us_xwalks__census__zcta2county_yearly__2012.parquet")
xwalk11 <- read_parquet("us_xwalks__census__zcta2county_yearly__2011.parquet")
xwalk10 <- read_parquet("us_xwalks__census__zcta2county_yearly__2010.parquet")
xwalk09 <- read_parquet("us_xwalks__census__zcta2county_yearly__2009.parquet")
xwalk08 <- read_parquet("us_xwalks__census__zcta2county_yearly__2008.parquet")
xwalk07 <- read_parquet("us_xwalks__census__zcta2county_yearly__2007.parquet")
xwalk06 <- read_parquet("us_xwalks__census__zcta2county_yearly__2006.parquet")
xwalk05 <- read_parquet("us_xwalks__census__zcta2county_yearly__2005.parquet")
xwalk04 <- read_parquet("us_xwalks__census__zcta2county_yearly__2004.parquet")
xwalk03 <- read_parquet("us_xwalks__census__zcta2county_yearly__2003.parquet")
xwalk02 <- read_parquet("us_xwalks__census__zcta2county_yearly__2002.parquet")
xwalk01 <- read_parquet("us_xwalks__census__zcta2county_yearly__2001.parquet")
xwalk00 <- read_parquet("us_xwalks__census__zcta2county_yearly__2000.parquet")
```

### Load data for total CO2 emissions (at census block group level, in kilograms)

```{r}
# Create path to the .gdb directory
co2_path <- "CMS_DARTE_V2_1735/data/DARTE_v2.gdb"

# List the available layers (feature classes) inside the .gdb
st_layers(dsn = co2_path)

# Load the specific layer
co2_cbg <- st_read(dsn = co2_path, layer = "DARTE_v2_blockgroup_kgco2_1980_2017")
```

### Load data for average PM2.5 concentration (at ZCTA level, in micrograms per cubic meter)

```{r}
# Create path to the .parquet files
pm25_path <- "dataverse_files"

# List all .parquet files for years 2000–2020
file_list <- list.files(path = pm25_path,
                        pattern = "^pm25__ushap__zcta_yearly__20\\d{2}\\.parquet$",
                        full.names = TRUE)

# Read and combine all files into data frame
pm25_zcta <- file_list %>%
  lapply(read_parquet) %>%
  bind_rows()
```

# Data wrangling for `co2_cbg`

### Convert `co2_cbg` to tidy format

```{r}
co2_cbg <- co2_cbg %>%
  pivot_longer(cols = starts_with("kgco2_"),
               names_to = "year",
               names_prefix = "kgco2_",
               values_to = "value") %>%
  mutate(year = as.integer(year))
```

<!-- ### Convert `co2_cbg` from kilograms to million metric tons (1 mmt = 1e9 kg) -->

<!-- ### Keep in kilograms moving forward -->

<!-- ```{r} -->
<!-- co2_cbg <- co2_cbg %>% -->
<!--   mutate(value = value / 1e9) -->
<!-- ``` -->

### Select for variables to avoid large data set

```{r}
co2_cbg <- co2_cbg %>%
  ungroup() %>%
  st_drop_geometry() %>%
  as.data.frame() %>%
  rename(geoid = GEOID) %>%
  select(geoid, year, value)
```

# Data wrangling for `xwalk`

### Combine everything into one crosswalk

```{r}
years <- 2000:2022
file_paths <- paste0("us_xwalks__census__zcta2county_yearly__", years, ".parquet")
xwalk_list <- map(file_paths, read_parquet)
xwalk <- bind_rows(xwalk_list)
```

# Aggregation for `pm25_zcta`

### Join `xwalk` and `pm25_zcta`

```{r}
pm25_county <- inner_join(xwalk, pm25_zcta, by = c("zcta", "year"))
```

### Aggregate from ZCTA level to county level

```{r}
pm25_county <- pm25_county %>% 
  group_by(year, county) %>% 
  mutate(weight = (arealand_pct / 100)) %>% 
  mutate(weight =  weight / sum(weight, na.rm = TRUE)) %>% 
  summarize(pm25 = sum(PM25 * weight, na.rm = TRUE),
            .groups = "drop")
```

# Aggregation for `co2_cbg`

### Create variable for county FIPS code

```{r}
co2_cbg <- co2_cbg %>%
  mutate(county = substring(geoid, 1, 5))
```

### Aggregate from census block group level to county level

```{r}
co2_county <- co2_cbg %>%
  group_by(year, county) %>% 
  summarize(co2 = sum(value, na.rm = TRUE),
            .groups = "drop")
```

# Finalizing files

### Filter for contiguous U.S.

```{r}
pm25_county <- pm25_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)

co2_county <- co2_county %>% 
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

### Create `df` by joining `pm25_county` and `co2_county`

```{r}
df <- inner_join(pm25_county, co2_county, by = c("county", "year")) %>%
  filter(substr(county, 1, 2) %in% contiguous_fips)
```

### Filter for years less than or equal to 2010 (to avoid confounding)

```{r}
df <- df %>%
  filter(year <= 2010)
```

# Data visualization

### Aggregate CO2

```{r}
# Aggregate total CO2 for all counties by year
co2_total_year <- co2_county %>%
  group_by(year) %>%
  summarize(total_co2 = sum(co2, na.rm = TRUE))

# Check the result
head(co2_total_year)
```

### Plot aggregate CO2

```{r}
# Create graph of CO2 data
co2_plot <- co2_total_year %>%
  ggplot() +
  geom_line(aes(x = year, y = total_co2), linewidth = 2) +
  geom_vline(aes(xintercept = 2005, linetype = "2005: Nonattainment Designation from NAAQS"), 
             color = "red",
             linewidth = 2.5) +
  labs(title = "Aggregated On-Road CO2 Emissions in the Contiguous U.S.",
       x = "Year",
       y = "CO2 (kilograms)",
       fill = "",
       linetype = "") +
  scale_linetype_manual(values = c("2005: Nonattainment Designation from NAAQS" = "dotted")) +
  theme_minimal() +
  # theme(legend.position = "bottom",
  #       plot.title = element_text(size = 44),
  #       axis.text = element_text(size = 32),
  #       axis.title = element_text(size = 40),
  #       legend.text = element_text(size = 36))
  theme(legend.position = "bottom")

# Plot and save graph
co2_plot
ggsave("co2_plot.png",
       plot = co2_plot,
       width = 20, height = 12, dpi = 300,
       bg = "white")
```

# Individual county-level analysis

### Run CausalArima on specified county to estimate statistics 

```{r}
# FIPS "01053" is in Alabama (which had minimal legislation besides NAAQS)

# Specify FIPS code for individual county
my_fips <- "01053"

# Get county_df for specified FIPS
county_df <- co2_county %>%
  filter(county == my_fips)

# Filter for 2010
county_df <- county_df %>%
  filter(year <= 2010)

# Define intervention time point
intervention_date <- as.Date("2005-01-01")

# Create vector for dates
all_dates <- as.Date(paste0(county_df$year, "-01-01"))

# Create time series for outcome with yearly seasonality
y_ts <- ts(county_df$co2 * 100000, frequency = 1)

# Run CausalArima()
ce <- CausalArima(y = y_ts,
                  dates = all_dates,
                  int.date = intervention_date,
                  nboot = 1000)

# calculate (observed - forecast) / forecast
# (y1 - y0) / y0
# reduced co2 with respect what would have been 

# calculate this year by year 

# extract forecasted series after running model with 2010
# Create list to store relative effects 


# Get impact as list
imp <- impact(ce)

# Extract statistics via impact_norm
norm_effect <- imp$impact_norm$sum
cum_effect <- norm_effect$estimate
avg_effect <- imp$impact_norm$average$estimate
sd_norm <- norm_effect$sd
ci_lower_norm <- cum_effect - 1.96 * sd_norm
ci_upper_norm <- cum_effect + 1.96 * sd_norm

# Extract statistics via impact_boot
boot_list <- imp$impact_boot
boot_effect <- boot_list$effect_cum[3, ]
rel_cum_effect <- imp$impact_boot$effect_cum$estimates[4]
p_value_boot_one_sided <- as.numeric(boot_list$p_values["p"])

?CausalArima

# Store data as tibble
object <- tibble(cum_effect = cum_effect,
                 rel_cum_effect = rel_cum_effect,
                 avg_effect = avg_effect,
                 sd_norm = sd_norm,
                 ci_lower_norm = ci_lower_norm,
                 ci_upper_norm = ci_upper_norm,
                 p_value_norm = norm_effect$p_value_bidirectional,
                 sd_boot = boot_effect$sd,
                 ci_lower_boot = boot_effect$inf,
                 ci_upper_boot = boot_effect$sup,
                 p_value_boot = min(1, 2 * p_value_boot_one_sided))
object

ce$dates
```

# County-level analysis across counties

### Run CausalArima on every available county to estimate statistics

```{r, cache = TRUE}
# Initialize list to store results
results_list <- list()

# Define intervention time point
intervention_date <- as.Date("2005-01-01")

# Get unique counties (to avoid repeating counties over years)
fips_vec <- unique(df$county)

# Loop through each county
for (i in seq_along(fips_vec)) {
  my_fips <- fips_vec[i]

  # Get county_df for specified FIPS
  county_df <- co2_county %>%
    filter(county == my_fips)

  # Skip if no data
  if (nrow(county_df) == 0) {
    warning(glue("No data for FIPS {my_fips}, skipping..."))
    next
  }

  # Create vector for dates
  all_dates <- as.Date(paste0(county_df$year, "-01-01"))

  # Create time series for outcome with yearly seasonality
  y_ts <- ts(county_df$co2, frequency = 1)

  # Use tryCatch to avoid full script crashing
  tryCatch({
    # Run CausalArima()
    ce <- CausalArima(y = y_ts,
                      dates = all_dates,
                      int.date = intervention_date,
                      nboot = 1000)

    # Get impact as list
    imp <- impact(ce)

    # Extract statistics via impact_norm
    norm_effect <- imp$impact_norm$sum
    cum_effect <- norm_effect$estimate
    avg_effect <- imp$impact_norm$average$estimate
    sd_norm <- norm_effect$sd
    ci_lower_norm <- cum_effect - 1.96 * sd_norm
    ci_upper_norm <- cum_effect + 1.96 * sd_norm

    # Extract statistics via impact_boot
    boot_list <- imp$impact_boot
    boot_effect <- boot_list$effect_cum[3, ]
    rel_cum_effect <- imp$impact_boot$effect_cum$estimates[4]
    p_value_boot_one_sided <- as.numeric(boot_list$p_values["p"])

    # Store data as tibble
    result_row <- tibble(fips = my_fips,
                         avg_effect = avg_effect,
                         cum_effect = cum_effect,
                         rel_cum_effect = rel_cum_effect,
                         sd_norm = sd_norm,
                         ci_lower_norm = ci_lower_norm,
                         ci_upper_norm = ci_upper_norm,
                         p_value_norm = norm_effect$p_value_bidirectional,
                         sd_boot = boot_effect$sd,
                         ci_lower_boot = boot_effect$inf,
                         ci_upper_boot = boot_effect$sup,
                         p_value_boot = min(1, 2 * p_value_boot_one_sided))

    # Add to current results
    results_list[[length(results_list) + 1]] <- result_row

    # Print progress
    message(glue("{i}/{length(fips_vec)}: FIPS {my_fips} processed."))

    # Save checkpoint every 50 iterations
    if (i %% 50 == 0 || i == length(fips_vec)) {
      checkpoint_df <- bind_rows(results_list)
      write_csv(checkpoint_df, "checkpoint_results.csv")
      message(glue("Checkpoint saved at iteration {i}"))
    }

  }, error = function(e) {
    warning(glue("Error for FIPS {my_fips}: {e$message}"))
  })
}

# Save and export results
df_causal_arima <- bind_rows(results_list)
message("All counties processed. Final results saved.")
```

### Export `df_causal_arima`

```{r}
write_csv(df_causal_arima, "~/df_causal_arima.csv")
```

### Read `df_causal_arima`

```{r}
df_causal_arima <- read_csv("df_causal_arima.csv")
```

### Create `df_map` by adding polygons to `df_causal_arima`

```{r}
# Join shapefile with df_causal_arima
df_map <- counties_sf %>%
 inner_join(df_causal_arima, by = "fips")
```

### Filter for significant counties (`p_value_boot` < 0.05)

```{r}
df_map_sig <- df_map %>% 
  filter(p_value_boot < 0.05)
```

### Plot choropleth map of U.S. with cumulative effect at each county, sequential

```{r}
# Define sequential color palette
seq_colors <- c("#021B40", "#1F78B4", "#D6ECF5")

# Compute global limits for cum_effect
global_min <- min(df_map_sig$cum_effect, na.rm = TRUE)
global_max <- max(df_map_sig$cum_effect, na.rm = TRUE)

# Create graph of cumulative effects
sig_cum_effect_map <- ggplot(data = df_map_sig) + 
  geom_sf(aes(fill = cum_effect), color = NA) + 
  scale_fill_gradientn(colors = seq_colors, 
                       na.value = "white", 
                       limits = c(global_min, 0), 
                       labels = label_number(accuracy = 1), 
                       guide = guide_colorbar(barwidth = 20, # 50
                                              barheight = 1, 
                                              title.position = "top", 
                                              title.vjust = 1)) + 
  labs(title = glue("Cumulative Effect by County, 2006-2010: p < 0.05"), 
       fill = "CO2 (kilograms)") + 
  theme_minimal() + 
  # theme(legend.position = "bottom",
  #       plot.title = element_text(size = 50),
  #       axis.text = element_text(size = 40),
  #       legend.title = element_text(size = 50),
  #       legend.text = element_text(size = 45)) +
  theme(legend.position = "bottom") +
  coord_sf(xlim = c(-125, -66), ylim = c(24, 50), expand = FALSE)
  
# Plot and save graph
sig_cum_effect_map
ggsave("sig_cum_effect_map.png",
       plot = sig_cum_effect_map,
       width = 20, height = 18, dpi = 300,
       bg = "white")
```

### Plot choropleth map of U.S. with cumulative effect at each county, diverging

```{r}
# Define diverging color palette
div_colors <- c("#2166AC", "#4393C3", "#92C5DE", 
                "#F7F7F7",
                "#F4A582", "#D6604D", "#B2182B")

# Compute global limits for cum_effect
global_min <- min(df_map_sig$cum_effect, na.rm = TRUE)
global_max <- max(df_map_sig$cum_effect, na.rm = TRUE)

# Create graph of cumulative effects
sig_cum_effect_map <- ggplot(data = df_map_sig) + 
  geom_sf(aes(fill = cum_effect), color = NA) +
  scale_fill_gradientn(colors = div_colors,
                       na.value = "white",
                       limits = c(global_min, global_max),
                       values = scales::rescale(c(global_min, 0, global_max)),
                       labels = label_number(accuracy = 1), 
                       guide = guide_colorbar(barwidth = 20, # 50
                                              barheight = 1, 
                                              title.position = "top", 
                                              title.vjust = 1)) + 
  labs(title = glue("Cumulative Effect by County, 2006-2010: p < 0.05"),
    fill = "CO2 (kilograms)") +
  theme_minimal() +
  # theme(legend.position = "bottom",
  #       plot.title = element_text(size = 50),
  #       axis.text = element_text(size = 40),
  #       legend.title = element_text(size = 50),
  #       legend.text = element_text(size = 45)) +
  theme(legend.position = "bottom") +
  coord_sf(xlim = c(-125, -66), ylim = c(24, 50), expand = FALSE)

# Plot and save graph
sig_cum_effect_map
ggsave("sig_cum_effect_map.png",
       plot = sig_cum_effect_map,
       width = 20, height = 18, dpi = 300,
       bg = "white")
```

# Data wrangling

### Transform `df` from county-year level to county level

```{r}
df <- df %>%
  select(county, year, pm25, co2) %>%
  pivot_wider(id_cols = county,
              names_from = year,
              values_from = c(pm25, co2),
              names_glue = "{.value}_{year}",
              values_fn = list(pm25 = ~first(na.omit(.)),
                               co2 = ~first(na.omit(.))),
              values_fill = NA_real_) %>%
  select(county,
         all_of(paste0("pm25_", 2000:2010)),
         all_of(paste0("co2_", 2000:2010)))
```

### Combine `df` with `df_causal_arima` (already at county level)

```{r}
df <- df %>%
  left_join(df_causal_arima %>% select(fips,
                                       rel_cum_effect,
                                       avg_effect,
                                       sd_boot,
                                       p_value_boot),
            by = c("county" = "fips"))
```

### Calculate 2005-2010 change in PM2.5 concentration for each county

```{r}
df <- df %>%
  mutate(pm25_change_2005_2010 = pm25_2010 - pm25_2005,
         pm25_rel_change_2005_2010 = pm25_change_2005_2010 / pm25_2005)
```

### Add variable for signficance (`p_value_boot` < 0.05)

```{r}
df <- df %>%
  mutate(significance = case_when(p_value_boot < 0.05 ~ 1,
                                  p_value_boot >= 0.05 ~ 0))
```

# Regression analysis (for counties with `significance` == 1)

### Filter for significant counties (`significance` == 1)

```{r}
df_sig <- df %>% 
  filter(significance == 1)
```

### Plot scatterplots

```{r}
# Visualize rel_cum_effect and pm25_2005
plot1 <- ggplot(df_sig, aes(y = rel_cum_effect,
                            x = pm25_2005)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize rel_cum_effect and pm25_rel_change_2005_2010
plot2 <- ggplot(df_sig, aes(y = rel_cum_effect,
                            x = pm25_rel_change_2005_2010)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)
```

### Run OLS linear regression with average effect as dependent variable 

```{r}
# Fit model
model_ols <- lm(avg_effect ~ pm25_2005 + pm25_rel_change_2005_2010,
                data = df_sig)

get_regression_table(model_ols)
```

### Plot scatterplots with OLS model

```{r}
# Visualize avg_effect and pm25_2005
plot1 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_2005)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize avg_effect and pm25_rel_change_2005_2010
plot2 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_rel_change_2005_2010)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)
```

### Run WLS linear regression with average effect as dependent variable 

```{r}
# Create weight from inverse variance
df_sig <- df_sig %>% 
  mutate(weight = 1 / (sd_boot)^2)

# Fit full model
model_wls <- lm(avg_effect ~ pm25_2005 + pm25_rel_change_2005_2010,
                weights = weight,
                data = df_sig)
get_regression_table(model_wls)
```

### Plot scatterplots with WLS model

```{r}
# Predicted values from the WLS model (holding other variable at mean)
newdata1 <- df_sig %>%
  mutate(pm25_rel_change_2005_2010 = mean(pm25_rel_change_2005_2010, na.rm = TRUE))
newdata1$pred <- predict(model_wls, newdata = newdata1)

# Plot with WLS model
plot1 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_2005,
                            size = weight)) +
  # ylim(-2000000000, 1000000000) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_line(data = newdata1, aes(y = pred, x = pm25_2005),
            color = "blue", linewidth = 1.1) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005–2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1

# Predicted values from the WLS model (holding other variable at mean)
newdata2 <- df_sig %>%
  mutate(pm25_2005 = mean(pm25_2005, na.rm = TRUE))
newdata2$pred <- predict(model_wls, newdata = newdata2)

# Plot with WLS model
plot2 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_rel_change_2005_2010,
                            size = weight)) +
  geom_point(alpha = 0.25) +
  # ylim(-2000000000, 1000000000) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_line(data = newdata2, aes(y = pred, x = pm25_rel_change_2005_2010),
            color = "blue", linewidth = 1.1) +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005–2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
```

### Fit smoothing splines

```{r}
# Fit model
gam_model <- gam(avg_effect ~ s(pm25_2005 + pm25_rel_change_2005_2010), weights = weight,
                 data = df_sig)

# Summarize model
summary(gam_model)
tidy(gam_model)

# Visualize model
plot(gam_model)
plot(gam_model, pages = 1, shade = TRUE, shade.col = "red",
     residuals = TRUE, pch = 16, cex = 0.1, scheme = 1, rug = TRUE)
```

### Plot scatterplots with smoothing splines

```{r}
# Visualize avg_effect and pm25_2005 with smoothing spline
plot1 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_2005)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(aes(x = pm25_2005, y = avg_effect, group = 1),
              method = "gam",
              formula = y ~ s(x, bs = "cs"),
              color = "blue",
              fill = "grey80",
              linewidth = 1.1,
              alpha = 0.2,
              inherit.aes = FALSE) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize avg_effect and pm25_rel_change_2005_2010 with smoothing spline
plot2 <- ggplot(df_sig, aes(y = avg_effect,
                            x = pm25_rel_change_2005_2010)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(aes(x = pm25_rel_change_2005_2010, y = avg_effect, group = 1),
              method = "gam",
              formula = y ~ s(x, bs = "cs"),
              color = "blue",
              fill = "grey80",
              linewidth = 1.1,
              alpha = 0.2,
              inherit.aes = FALSE) + 
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)

# Visualize pm25_2005 and pm25_change_2005_2010
ggplot(df_sig, aes(y = pm25_2005, x = pm25_rel_change_2005_2010)) +
  geom_point()
```

# Regression analysis (for all counties)

### Plot scatterplots

```{r}
# Visualize rel_cum_effect and pm25_2005
plot1 <- ggplot(df, aes(y = rel_cum_effect,
                        x = pm25_2005,
                        color = factor(significance))) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize rel_cum_effect and pm25_rel_change_2005_2010
plot2 <- ggplot(df, aes(y = rel_cum_effect,
                        x = pm25_rel_change_2005_2010,
                        color = factor(significance))) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)
```

### Run OLS linear regression with average effect as dependent variable 

```{r}
# Fit model
model_ols <- lm(avg_effect ~ pm25_2005 + pm25_rel_change_2005_2010,
                data = df)

get_regression_table(model_ols)
```

### Plot scatterplots with OLS model

```{r}
# Visualize avg_effect and pm25_2005
plot1 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_2005,
                        color = factor(significance))) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize avg_effect and pm25_rel_change_2005_2010
plot2 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_rel_change_2005_2010,
                        color = factor(significance))) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE) +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)
```

### Run WLS linear regression with average effect as dependent variable 

```{r}
# Create weight from inverse variance
df <- df %>% 
  mutate(weight = 1 / (sd_boot)^2)

# Fit full model
model_wls <- lm(avg_effect ~ pm25_2005 + pm25_rel_change_2005_2010,
                weights = weight,
                data = df)
get_regression_table(model_wls)
```

### Plot scatterplots with WLS model

```{r}
# Predicted values from the WLS model (holding other variable at mean)
newdata1 <- df %>%
  mutate(pm25_rel_change_2005_2010 = mean(pm25_rel_change_2005_2010, na.rm = TRUE))
newdata1$pred <- predict(model_wls, newdata = newdata1)

# Plot with WLS model
plot1 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_2005,
                        color = factor(significance),
                        size = weight)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_line(data = newdata1, aes(y = pred, x = pm25_2005),
            color = "blue", linewidth = 1.1) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005–2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1

# Predicted values from the WLS model (holding other variable at mean)
newdata2 <- df %>%
  mutate(pm25_2005 = mean(pm25_2005, na.rm = TRUE))
newdata2$pred <- predict(model_wls, newdata = newdata2)

# Plot with WLS model
plot2 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_rel_change_2005_2010,
                        color = factor(significance),
                        size = weight)) +
  geom_point(alpha = 0.25) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_line(data = newdata2, aes(y = pred, x = pm25_rel_change_2005_2010),
            color = "blue", linewidth = 1.1) +
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       size = "Weight (estimated inverse variance)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005–2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
```

### Fit smoothing splines

```{r}
# Fit model
gam_model <- gam(avg_effect ~ s(pm25_2005 + pm25_rel_change_2005_2010), weights = weight,
                 data = df)

# Summarize model
summary(gam_model)
tidy(gam_model)

# Visualize model
plot(gam_model)
plot(gam_model, pages = 1, shade = TRUE, shade.col = "red",
     residuals = TRUE, pch = 16, cex = 0.1, scheme = 1, rug = TRUE)
```

### Plot scatterplots with smoothing splines

```{r}
# Visualize avg_effect and pm25_2005 with smoothing spline
plot1 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_2005,
                        color = factor(significance))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(aes(x = pm25_2005, y = avg_effect, group = 1),
              method = "gam",
              formula = y ~ s(x, bs = "cs"),
              color = "black",
              fill = "grey80",
              linewidth = 1.1,
              alpha = 0.2,
              inherit.aes = FALSE) +
  labs(x = "PM2.5 Concentration in 2005 (micrograms per cubic meter)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       title = "Starting PM2.5 Concentration and ATE on CO2 Emissions, 2005-2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8))
plot1
ggsave("plot1.png", plot = plot1)

# Visualize avg_effect and pm25_rel_change_2005_2010 with smoothing spline
plot2 <- ggplot(df, aes(y = avg_effect,
                        x = pm25_rel_change_2005_2010,
                        color = factor(significance))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(aes(x = pm25_rel_change_2005_2010, y = avg_effect, group = 1),
              method = "gam",
              formula = y ~ s(x, bs = "cs"),
              color = "black",
              fill = "grey80",
              linewidth = 1.1,
              alpha = 0.2,
              inherit.aes = FALSE) + 
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       color = "Significance",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)

# Visualize pm25_2005 and pm25_change_2005_2010
ggplot(df, aes(y = pm25_2005, x = pm25_rel_change_2005_2010)) +
  geom_point()
```

### Model and visualize with only counties that reduced PM2.5 from 2005 to 2010

```{r}
# Filter for counties that reduced PM2.5 from 2005 to 2010
df_red <- df %>% 
  filter(pm25_change_2005_2010 < 0) %>% 
  mutate(weight = 1 / (sd_boot)^2)

# Visualize avg_effect and pm25_rel_change_2005_2010 with OLS model
plot2 <- ggplot(df_red, aes(y = avg_effect,
                            x = pm25_rel_change_2005_2010)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm") + 
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)

# Fit model with OLS
model_test_ols <- lm(avg_effect ~ pm25_rel_change_2005_2010, data = df_red)
get_regression_table(model_test_ols)

# Fit model with WLS
model_test_wls <- lm(avg_effect ~ pm25_rel_change_2005_2010, weights = weight, data = df_red)
get_regression_table(model_test_wls)

# Fit two-predictor model with OLS
model_test_two_ols <- lm(avg_effect ~ pm25_rel_change_2005_2010 + pm25_2005,
                         data = df_red)
get_regression_table(model_test_two_ols)

# Fit two-predictor model with WLS
model_test_two_wls <- lm(avg_effect ~ pm25_rel_change_2005_2010 + pm25_2005,
                         weights = weight,
                         data = df_red)
get_regression_table(model_test_two_wls)
```

### Model and visualize with only significant counties that reduced PM2.5 from 2005 to 2010

```{r}
# Filter for significant counties that reduced PM2.5 from 2005 to 2010
df_red_sig <- df %>% 
  filter(pm25_change_2005_2010 < 0, significance == 1) %>% 
  mutate(weight = 1 / (sd_boot)^2)

# Visualize avg_effect and pm25_rel_change_2005_2010 with OLS model
plot2 <- ggplot(df_red_sig, aes(y = avg_effect,
                                x = pm25_rel_change_2005_2010)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, color = "black", linetype = "dashed") +
  geom_vline(xintercept = 0, color = "black", linetype = "dashed") +
  geom_smooth(method = "lm") + 
  labs(x = "PM2.5 Relative Change from 2005 to 2010 (%)",
       y = "Estimated ATE of NAAQS on CO2 Emissions (kilograms)",
       title = "PM2.5 Change and ATE on CO2 Emissions, 2005-2010",
       caption = "A positive x-value means PM2.5 increased from 2005 to 2010") +
  theme(plot.title = element_text(size = 12),
        axis.title.x = element_text(size = 8),
        axis.title.y = element_text(size = 8),
        axis.text = element_text(size = 8),
        plot.caption = element_text(size = 8))
plot2
ggsave("plot2.png", plot = plot2)

# Fit model with OLS
model_test_ols <- lm(avg_effect ~ pm25_rel_change_2005_2010, data = df_red_sig)
get_regression_table(model_test_ols)

# Fit model with WLS
model_test_wls <- lm(avg_effect ~ pm25_rel_change_2005_2010, weights = weight, data = df_red_sig)
get_regression_table(model_test_wls)

# Fit two-predictor model with OLS
model_test_two_ols <- lm(avg_effect ~ pm25_rel_change_2005_2010 + pm25_2005,
                         data = df_red_sig)
get_regression_table(model_test_two_ols)

# Fit two-predictor model with WLS
model_test_two_wls <- lm(avg_effect ~ pm25_rel_change_2005_2010 + pm25_2005,
                         weights = weight,
                         data = df_red_sig)
get_regression_table(model_test_two_wls)
```
